### 2019-09-29 script/RNotebook/analysis.Rmd
Add IOI analyses
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

##### Description: This is a summary of all analyses for "The Sound of Teaching Music 1".

##### R Notebook:
- Execute a chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.
- Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.
- When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 
- The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

# Outline

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
# statistics
if (!require("stats")) {install.packages("stats"); require("stats")}
if (!require("ez")) {install.packages("ez"); require("ez")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "top")
```

# Interonset intervals (IOIs)
```{r extract, include = FALSE}
df_ioi <- read.csv("./trimmed/data_ioi.csv", header = T, sep = ",", dec = ".") # read a trimmed csv

# SubNr as a factor
df_ioi$SubNr <- as.factor(df_ioi$SubNr)

# Include only articulation
df_ioi_art <- df_ioi %>% dplyr::filter(Skill == "articulation")
# Include only dynamics
df_ioi_dyn <- df_ioi %>% dplyr::filter(Skill == "dynamics")
```

## Articulation
### 1. All - individual
```{r ioi_art, echo = FALSE}
# For each individual
ioi_art <- aggregate(IOI~SubNr*Condition*Skill, data = df_ioi_art,
                 FUN = function(x){c(length(x), mean = mean(x), sd = sd(x))})
ioi_art <- cbind(ioi_art[,1:3], as.data.frame(ioi_art[,4]))
# Change colnames
colnames(ioi_art) <- c("SubNr", "Condition", "Skill", "N", "Mean", "SD")
print(ioi_art)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_art_bar,  echo = FALSE}
p_ioi_art <- ggplot(data = ioi_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)")
plot(p_ioi_art)
```

### 2. All - group
```{r ioi_art_all, echo = FALSE}
# Group mean
ioi_art_all <- aggregate(Mean~Condition*Skill, data = ioi_art,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_art_all <- cbind(ioi_art_all[,1:2], as.data.frame(ioi_art_all[,3]))
# Change colnames
colnames(ioi_art_all) <- c("Condition", "Skill", "N", "Mean", "SD", "SEM")
print(ioi_art_all)
```

#### Plot
##### Bar chart
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_art_all_bar, echo = FALSE}
p_ioi_art_all <- ggplot(data = ioi_art_all, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)") + coord_cartesian(ylim = c(120, 210)) +
  theme(legend.position = "None")
p_ioi_art_all
```

##### Boxplot
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_art_all_box, echo = FALSE}
p_ioi_art_all_box <- ggplot(data = ioi_art, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)") + #coord_cartesian(ylim = c(100, 210)) +
  theme(legend.position = "None")
p_ioi_art_all_box
```

#### Statistics
##### Normality check
```{r normality1, echo = FALSE}
# Normality check
ioi_art_norm <- by(ioi_art$Mean, list(ioi_art$Condition), shapiro.test)
print(ioi_art_norm)
```

##### Two-way ANOVA (ez)
```{r stats1, echo = FALSE, message = FALSE}
ioi_art_aov <- ezANOVA(
  data = df_ioi_art
  , dv = .(IOI)
  , wid = .(SubNr)
  , within = .(Condition)
  , type = 3
  , detailed = TRUE
)
print(ioi_art_aov)
```

##### Two-way ANOVA (aov)
```{r stats2, echo = FALSE, message = FALSE}
ioi_art_aov_2 <- aov(Mean ~ Condition + Error(SubNr/Condition), data = ioi_art)
print(summary(ioi_art_aov_2))
```

### 3. Phrase boundaries - individual
```{r ioi_art_ch, echo = FALSE}
# Define phrase boundaries
df_ioi_art$Boundary <- "No"
df_ioi_art$Boundary[df_ioi_art$Interval == 8 | df_ioi_art$Interval == 16 | df_ioi_art$Interval == 24 | df_ioi_art$Interval == 41 | df_ioi_art$Interval == 49 | df_ioi_art$Interval == 57] <- "Yes"

# For each individual
ioi_art_ch <- aggregate(IOI~SubNr*Condition*Skill*Boundary, data = df_ioi_art,
                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_art_ch <- cbind(ioi_art_ch[,1:4], as.data.frame(ioi_art_ch[,5]))
# Change colnames
colnames(ioi_art_ch) <- c("SubNr", "Condition", "Skill", "Boundary", "N", "Mean", "SD")
print(ioi_art_ch)
```

#### Plot
##### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_art_ch_bar, echo = FALSE}
p_ioi_art_ch <- ggplot(data = ioi_art_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Boundary ~ .) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 18, 188, label = "Tempo (80bpm)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)") + coord_cartesian(ylim = c(100, 280))
p_ioi_art_ch
```

### 4. Phrase boundaries - group
```{r ioi_art_ch_all, echo = FALSE}
# Group mean
ioi_art_ch_all <- aggregate(Mean~Condition*Skill*Boundary, data = ioi_art_ch,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_art_ch_all <- cbind(ioi_art_ch_all[,1:3], as.data.frame(ioi_art_ch_all[,4]))
# Change colnames
colnames(ioi_art_ch_all) <- c("Condition", "Skill", "Boundary", "N", "Mean", "SD", "SEM")
print(ioi_art_ch_all)
```

#### Plot
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_art_ch_all_bar, echo = FALSE}
p_ioi_art_ch_all_bar <- ggplot(data = ioi_art_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1, 188, label = "Tempo (80bpm)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  #facet_grid(. ~ Boundary) +
  labs(y = "IOI (ms)") + coord_cartesian(ylim = c(100, 250))
p_ioi_art_ch_all_bar
```

##### Boxplot
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_art_ch_all_box, echo = FALSE}
p_ioi_art_ch_box <- ggplot(data = ioi_art_ch, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y" ,stackdir = "center", dotsize = 1, position=position_dodge(0.75)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)")
p_ioi_art_ch_box
```

#### Statistics
##### Normality check
```{r norrmality2, echo = FALSE}
ioi_art_ch_norm <- by(ioi_art_ch$Mean, list(ioi_art_ch$Condition, ioi_art_ch$Boundary), shapiro.test)
print(ioi_art_ch_norm)
```

##### Two-way ANOVA (ez)
```{r stats3, echo = FALSE}
ioi_art_ch_aov <- ezANOVA(
  data = df_ioi_art
  , dv = .(IOI)
  , wid = .(SubNr)
  , within = .(Condition, Boundary)
  , type = 3
  , detailed = TRUE
)
print(ioi_art_ch_aov)
```

##### Two-way ANOVA (aov)
```{r stats4, echo = FALSE}
ioi_art_ch_aov_2 <- with(ioi_art_ch,
  aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary))))
print(summary(ioi_art_ch_aov_2))
```

### 5. Trial by trial - individual
```{r ioi_art_trial, echo = FALSE}
# For each individual
ioi_art_trial <- aggregate(IOI~SubNr*Condition*Skill*TrialNr, data = df_ioi_art,
                    FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
ioi_art_trial <- cbind(ioi_art_trial[,1:4], as.data.frame(ioi_art_trial[,5]))
# Change colnames
colnames(ioi_art_trial) <- c("SubNr", "Condition", "Skill", "TrialNr", "N", "Mean", "SD")
print(ioi_art_trial)
```

#### Plot
##### Line graph
- A dashed line represents the tempo given by a metronome (80bpm)
```{r ioi_art_trial_line, echo = FALSE, fig.height = 8}
p_ioi_art_trial_line <- ggplot(data = ioi_art_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "IOI (ms)") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_art_trial_line
```

### 6. Trial by trial - group
```{r ioi_art_trial_all, echo = FALSE}
# Group mean
ioi_art_trial_all <- aggregate(Mean~Condition*Skill*TrialNr, data = ioi_art_trial,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
ioi_art_trial_all <- cbind(ioi_art_trial_all[,1:3], as.data.frame(ioi_art_trial_all[,4]))
# Change colnames
colnames(ioi_art_trial_all) <- c("Condition", "Skill", "TrialNr", "N", "Mean", "SD", "SEM")
print(ioi_art_trial_all)
```

#### Plot
##### Line graph
```{r ioi_art_trial_all_line, echo = FALSE}
p_ioi_art_trial_line <- ggplot(data = ioi_art_trial_all, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.05)) + 
  labs(x = "Trial", y = "IOI (ms)") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_art_trial_line
```

### 7. Variability all - individual

### 8. Variability all - group

### 9. Variability trial-by-trial - individual

### 10. Variability trial-by-trial - group

### Summary (IOI:articulation)
#### 1. All - individual
Some participants exhibited a strong tendency for slow demonstration while teaching.

#### 2. All - group
Participants played slower while teaching.

#### 3. Phrase boundaries - individual
Some participants played slower particularly at phrase boundaries.

#### 4. Phrase boundaries - group
Participants played slower while teaching / at phrase boundaries. However, there was no interaction.

#### 5. Trial by trial - individual
#### 6. Trial by trial - group
#### 7. Variability all - individual
#### 8. Variability all - group
#### 9. Variability trial-by-trial - individual
#### 10. Variability trial-by-trial - group