### 2019-10-05 script/RNotebook/questionnaire.Rmd
add summary for a questionnaire
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

##### Description: This is a summary of demographic information

# Outline - Quesionnaire

```{r read, echo = FALSE}
# Read csv files
df <- read.csv("./questionnaire.csv", header = T, sep = ",", dec = ".")
df_exc <- read.csv("./filtered/data_errorRate.csv", header = T, sep = ",", dec = ".")

# Change colnames
colnames(df) <- c("TimeStamp", "SubNr", "Age", "Gender", "FirstLang", "SecondLang", "DominantHand", "NormalVision", "NormalHearing", "Desc1", "Desc2",
                  "Instruments1", "FormalTraining1", "TotalPractice", "Start1", "End1", "PracticeHours", # Experience in musical instruments
                  "Teaching", "Instruments2", "TotalTeaching1", "FormalTraining2", "Training", "AgeGroup", "Start2", "End2", "Like1", # Experience in teaching instruments
                  "Other", "Subject", "TotalTeaching2", "Like2") # Experinence in teaching in general

# Extract the total number of years of training in piano (input manually based on the original csv file)
df$PianoTotalPractice <- c(14, 20, 27, 21, 10, 16, 9, 13, 25, 14, 13, 11, 6, 4, 6, 8, 21, 15, 18, 7, 11, 12, 12, 18, 12, 5, 5, 10, 16, 10, 8, 22, 10, 5, 4, 14)

print(df)
```

### Descriptive stats for participants' demographics
#### Number of excluded participants
```{r excluded, echo = FALSE}
# Descriptive statistics for participants' demographics
total <- nrow(df)
# Number of excluded participants
exp_error <- length(c(8)) # SubNr 8 - due to an experimental error
pitch_error <- length(df_exc$SubNr[df_exc$LessThan10 == "exclude"])
tempo_dev <- total-exp_error-pitch_error-length(include)
total_exc <- exp_error+pitch_error+tempo_dev

excluded_participants <- data.frame(cbind(total, total_exc, exp_error, pitch_error, tempo_dev))
colnames(excluded_participants) <- c("Total participants", "Total excluded", "Experimental error", "Pitch error", "Tempo deviation")
print(excluded_participants)
```

#### Number of included participants
```{r included, echo = FALSE}
# Data frame with only included participants
include <- df_exc$SubNr[df_exc$Final == "include"]
df_demo <- data.frame()
for (subnr in include){
  df_current <- df %>% dplyr::filter(SubNr == subnr)
  df_demo <- rbind(df_demo, df_current)
}

# Number of included participants
included <- nrow(df_demo)
male <- nrow(subset(df_demo, df_demo$Gender == "Male"))
female <- nrow(subset(df_demo, df_demo$Gender == "Female"))
experience_m <- round(mean(df_demo$PianoTotalPractice), 0)
experience_sd <- round(sd(df_demo$PianoTotalPractice)*sqrt((length(df_demo$PianoTotalPractice)-1)/length(df_demo$PianoTotalPractice)), 2) #biased sd for the current data

included_participants <- data.frame(cbind(total, included, male, female, experience_m, experience_sd))
colnames(included_participants) <- c("Total", "Included participants", "Male", "Female", "Average experience", "SD of Average experience")
print(included_participants)
```

#### Error Rate of excluded participants
```{r errorRate_excluded, echo = FALSE}
# Information about excluded participants due to errorRate (SubNr 16 was excluded but due to not errorRate but tempo deviation)
errorRate_mean_exc <- mean(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])*100
errorRate_sd_exc <- sd(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])*sqrt((length(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])-1)/length(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16]))*100 #biased sd for the current data
errorRate_median_exc <- median(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])*100
errorRate_min_exc <- min(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])*100
errorRate_max_exc <- max(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])*100

errorRate_excluded <- data.frame(cbind(errorRate_mean_exc, errorRate_sd_exc, errorRate_median_exc, errorRate_min_exc, errorRate_max_exc))
colnames(errorRate_excluded) <- c("ER Mean", "ER SD", "ER Median", "ER Min", "ER Max")
print(errorRate_excluded)
```

#### Error Rate: Normality test for excluded participants
```{r normality_for_excluded, echo = FALSE}
# Normality check
errorRate_excluded_norm <- shapiro.test(df_exc$ErrorRate[df_exc$Final == "exclude" & df_exc$SubNr != 16])
print(errorRate_excluded_norm)

# Note: The distributions fo error rates were not normal, so we decided to report median values (plus its range) rather than mean values.
```

#### Error Rate of included participants
```{r er_included, echo = FALSE}
# Information about included participants
errorRate_mean_inc <- mean(df_exc$ErrorRate[df_exc$Final == "include"])*100
errorRate_sd_inc <- sd(df_exc$ErrorRate[df_exc$Final == "include"])*sqrt((length(df_exc$ErrorRate[df_exc$Final == "include"])-1)/length(df_exc$ErrorRate[df_exc$Final == "include"]))*100 #biased sd for the current data
errorRate_median_inc <- median(df_exc$ErrorRate[df_exc$Final == "include"])*100
errorRate_min_inc <- min(df_exc$ErrorRate[df_exc$Final == "include"])*100
errorRate_max_inc <- max(df_exc$ErrorRate[df_exc$Final == "include"])*100

errorRate_included <- data.frame(cbind(errorRate_mean_inc, errorRate_sd_inc, errorRate_median_inc, errorRate_min_inc, errorRate_max_inc))
colnames(errorRate_included) <- c("ER Mean", "ER SD", "ER Median", "ER Min", "ER Max")
print(errorRate_included)
```

#### Error Rate: Normality test for included participants
```{r normality_for_included, echo = FALSE}
# Normality check
errorRate_included_norm <- shapiro.test(df_exc$ErrorRate[df_exc$Final == "include"])
print(errorRate_included_norm)

# Note: The distributions fo error rates were not normal, so we decided to report median values (plus its range) rather than mean values.
```