### 2020-04-02 script/RNotebook/additional020420/articulation.Rmd
additional analysis for articulation transitions
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Additional analysis for transition points"
output: html_notebook
---
Description: These are some additional analyses for transition points.

## Articulation
How does it look like articulation transition points?
Posthoc prediciton: Participants would have a larger separation betwenn the end of legato and the beginning of staccato.

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
# statistics
if (!require("stats")) {install.packages("stats"); require("stats")}
if (!require("effsize")) {install.packages("effsize"); require("effsize")}
if (!require("ez")) {install.packages("ez"); require("ez")}

# read csv
df_kot <- fread("./../trimmed/data_kot.csv")

# only transition points for arituclation
df_kot_trans <- df_kot[Skill == "articulation" & Subcomponent != "Legato" & Subcomponent != "Staccato"]
```

```{r stats1, echo = TRUE}
# For each individual
art_trans <- df_kot_trans[, .(N = length(KOT), Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Subcomponent)]
print(art_trans)
# Group
art_trans_all <- art_trans[, .(N = length(Mean), Mean = mean(Mean), SEM = sd(Mean)/sqrt(length(Mean))), by = .(Condition, Subcomponent)]
print(art_trans_all)
# Stats
# normality check
art_trans_norm <- by(art_trans$Mean, list(art_trans$Condition, art_trans$Subcomponent), shapiro.test)
print(art_trans_norm)
# ezAnova
art_trans_aov <- ezANOVA(
  data = df_kot_trans
  , dv = .(KOT)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
print(art_trans_aov)
```

### Plot
#### Bar chart
- Error bars show SEM of the mean

```{r art_trans_bar, echo = FALSE}　
p_art_trans <- ggplot(data = art_trans_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation Transition")
p_art_trans
```

#### Box plot
```{r art_trans_box, echo = FALSE}
p_art_trans_box <- ggplot(data = art_trans, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation Transition")
p_art_trans_box
```