### 2020-06-02 script/RNotebook/questionnaire.Rmd
upload analysis for questionnaire
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Demographics: The Sound of Teaching Music 2'
output:
  html_notebook: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

Description: This is a summary of demographic information

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
```

# Outline - Questionnaire

```{r read, echo = FALSE}
# Read csv files
df <- read.csv("./questionnaire_teaching-v2.0.csv", header = T, sep = ",", dec = ".", na.string = "NA")
df_error_onset <- read.csv("../R/preprocessor/filtered/data_error_correction_onset.csv", header = T, sep = ",", dec = ".")
df_error_offset <- read.csv("../R/preprocessor/filtered/data_error_correction_offset.csv", header = T, sep = ",", dec = ".")
df$Please.specify.the.total.number.of.years.of.formal.training..i.e...lessons.. <- as.numeric(df$Please.specify.the.total.number.of.years.of.formal.training..i.e...lessons..)

# add SubNr 21 manually
df[21,2] <- 21
df[21,3] <- 25
df[21,4] <- "Male"
df[21,5] <- "Hungarian"
df[21,6] <- "English"
df[21,7] <- "Right"
df[21,8] <- "Yes"
df[21,9] <- "Yes"
df[21,17] <- 11
df[21,18] <- 12
df[21,19] <- 13

# exclude SubNr 3 due to insufficient motor skills
include <- df %>% dplyr::filter(Participant.number != 3)

# colnames
colnames(include) <- c("TimeStamps", "SubNr", "Age", "Gender", "FirstLang", "SecondLang", "DominantHand",
                       "Vision", "Hearing", "KnowClementi", "PlayClementi", colnames(include)[12],
                       colnames(include)[13], "School", "Degree", "Major", "TotalTraining", "TotalPlaying",
                       "AgeStarted", "AgeStopped", "PracticeHours", "Instruments", colnames(include)[23],
                       colnames(include)[24], colnames(include)[25], colnames(include)[26],
                       colnames(include)[27], colnames(include)[28], colnames(include)[29],
                       colnames(include)[30], colnames(include)[31], colnames(include)[32],
                       colnames(include)[33], colnames(include)[34], colnames(include)[35],
                       "MusicTeachingExperience", "OtherTeachingExperience", "Subject", "TotalTeaching",
                       "TeachingLiking")

# correct some gibbrish
include[13,17] <- 18
print(include)
```

## Descriptive stats for participants' demographics
- Number of excluded participants: `r nrow(df)-nrow(include)`
- Number of included participants: `r nrow(include)`
- Male: `r nrow(subset(include, include$Gender == "Male"))`
- Female: `r nrow(subset(include, include$Gender == "Female"))`
- Dominant hand: Right - `r nrow(subset(include, include$DominantHand == "Right"))`, Left - `r nrow(subset(include, include$DominantHand == "Left"))`
- Average: the total number of years of formal learning: `r mean(include$TotalTraining)`
- Average: the total number of years of playing: `r mean(include$TotalPlaying)`