### 2020-07-29 script/RNotebook/additional240720/dynamics.Rmd
upload analysis with new trimmed
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Dynamics Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of all analyses for dynamics. **Add excluded 14 participants.**

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Outline - Dynamics (Key Velocity)
## [Experimental Design](#design)

## [Dynamics](#vel-dynamics)

### KV
- Individual results
  + [All](#vel_dyn_ind_all)
  + [Trial by trial](#vel_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#vel_dyn_group_all)
  + [Trial by trial](#vel_dyn_group_trial)

<!-- - Order analysis (teaching/performing first?) -->
<!--   + [All](#vel_dyn_order_group_all) -->

### KV-Diff
- Individual results
  + [All](#vel_acc_dyn_ind_all)
  + [Trial by trial](#vel_acc_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#vel_acc_dyn_group_all)
  + [Trial by trial](#vel_acc_dyn_group_trial)

<!-- - Order analysis (teaching/performing first?) -->
<!--   + [All](#vel_acc_dyn_order_group_all) -->

- [Sequence plot](#sequence_dyn)

## [Articulation](#vel_articulation)

### KV
- Individual results
  + [All](#vel_art_ind_all)
  + [Trial by trial](#vel_art_ind_trial)

- Group results and stats (if available)
  + [All](#vel_art_group_all)
  + [Trial by trial](#vel_art_group_trial)

<!-- - Order analysis (teaching/performing first?) -->
<!--   + [All](#vel_art_order_group_all) -->

### KV-Diff
- Individual results
  + [All](#vel_acc_art_ind_all)
  + [Trial by trial](#vel_acc_art_ind_trial)

- Group results and stats (if available)
  + [All](#vel_acc_art_group_all)
  + [Trial by trial](#vel_acc_art_group_trial)

<!-- - Order analysis (teaching/performing first?) -->
<!--   + [All](#vel_acc_art_order_group_all) -->

- [Sequence plot](#sequence_art)

## Experimental Design{#design}
- Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: forte vs. piano / legato vs. staccato)
- Dependent variables: Key Velocity (KV), KV difference

## Main predictions
Participants will exaggerate forte and piano more while teaching.

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
# statistics
if (!require("stats")) {install.packages("stats"); require("stats")}
if (!require("effsize")) {install.packages("effsize"); require("effsize")}
if (!require("ez")) {install.packages("ez"); require("ez")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r file, include = FALSE}
filename_vel = "./trimmed/data_vel.csv"
filename_vel_acc = "./trimmed/data_vel_acc.csv"
```

```{r extract, include = FALSE}
df_vel <- read.csv(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed csv
df_vel_acc <- read.csv(filename_vel_acc, header = T, sep = ",", dec = ".") # read a trimmed csv

# SubNr as a factor
df_vel$SubNr <- as.factor(df_vel$SubNr)
df_vel_acc$SubNr <- as.factor(df_vel_acc$SubNr)

# Include only articulation
df_vel_art <- df_vel %>% dplyr::filter(Skill == "articulation")
df_vel_acc_art <- df_vel_acc %>% dplyr::filter(Skill == "articulation")

# Include only dynamics
df_vel_dyn <- df_vel %>% dplyr::filter(Skill == "dynamics")
df_vel_acc_dyn <- df_vel_acc %>% dplyr::filter(Skill == "dynamics")
```

# Key Velocity (KV) - Dynamics
## 1. All - individual{#vel_dyn_ind_all}
- Average KV for each participant.

```{r vel_dyn, echo = FALSE}
# For each individual
vel_dyn <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent, data = df_vel_dyn,
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_dyn <- cbind(vel_dyn[,1:4], as.data.frame(vel_dyn[,5]))
# Change colnames
colnames(vel_dyn) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(vel_dyn)
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r vel_dyn_bar,  echo = FALSE}
p_vel_dyn <- ggplot(data = vel_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100))
p_vel_dyn
```

## 2. All - group{#vel_dyn_group_all}
- Average KV for each condition.
- Hypothesis: Participants will exaggerate forte and piano when teaching.

```{r vel_dyn_all, echo = FALSE}
# Group mean
vel_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = subset(vel_dyn, vel_dyn$SubNr != 3 & vel_dyn$SubNr != 14), # exclude SubNr 3 & 14 as one of the conditions is completely missing
                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_dyn_all <- cbind(vel_dyn_all[,1:3], as.data.frame(vel_dyn_all[,4]))
# Change colnames
colnames(vel_dyn_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
print(vel_dyn_all)
```

### Plot
#### Bar chart
- Error bars show SEM of the mean

```{r vel_dyn_all_bar, echo = FALSE}　
p_vel_dyn_all <- ggplot(data = vel_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100))
p_vel_dyn_all
```

#### Box plot
```{r vel_dyn_all_box, echo = FALSE}
p_vel_dyn_all_box <- ggplot(data = subset(vel_dyn, vel_dyn$SubNr != 3 & vel_dyn$SubNr != 14), aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KV (0-127)", subtitle = "KV: Dynamics")
p_vel_dyn_all_box
```

### Statistics
#### Normality check
```{r normality1, echo = FALSE}
# exclude SubNr 3 & 14
vel_dyn <- subset(vel_dyn, vel_dyn$SubNr != 3 & vel_dyn$SubNr != 14)
# Reduce unused factors
vel_dyn$Subcomponent <- factor(vel_dyn$Subcomponent)
vel_dyn_norm <- by(vel_dyn$Mean, list(vel_dyn$Condition, vel_dyn$Subcomponent), shapiro.test)
print(vel_dyn_norm)
```

#### Two-way ANOVA (ez)
```{r stats1, echo = FALSE}
vel_dyn_aov <- ezANOVA(
  data = subset(df_vel_dyn, df_vel_dyn$SubNr != 3 & df_vel_dyn$SubNr != 14)
  , dv = .(Velocity)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_dyn_aov$ANOVA$F <- round(vel_dyn_aov$ANOVA$F, 4)
vel_dyn_aov$ANOVA$p <- round(vel_dyn_aov$ANOVA$p, 4)
print(vel_dyn_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats2, echo = FALSE}
vel_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_dyn)
print(summary(vel_dyn_aov_2))
```

#### Post-hoc (direct comparison)
```{r post-hoc1, echo = FALSE}
vel_ttest_forte <- t.test(vel_dyn$Mean[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Forte"], vel_dyn$Mean[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Forte"], paired = TRUE)
vel_ttest_piano <- t.test(vel_dyn$Mean[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Piano"], vel_dyn$Mean[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Piano"], paired = TRUE)
# Effect size
vel_ttest_forte_cohend <- cohen.d(vel_dyn$Mean[vel_dyn$Subcomponent == "Forte"], vel_dyn$Condition[vel_dyn$Subcomponent == "Forte"], paired = TRUE)
vel_ttest_piano_cohend <- cohen.d(vel_dyn$Mean[vel_dyn$Subcomponent == "Piano"], vel_dyn$Condition[vel_dyn$Subcomponent == "Piano"], paired = TRUE)

print(vel_ttest_forte)
print(vel_ttest_piano)
print(vel_ttest_forte_cohend)
print(vel_ttest_piano_cohend)
```

#### Post-hoc (pairwise t test)
```{r post-hoc2, echo = FALSE}
# Create a group label for a post-hoc test
vel_dyn$Group[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Forte"] <- "Performing-Forte" 
vel_dyn$Group[vel_dyn$Condition == "performing" & vel_dyn$Subcomponent == "Piano"] <- "Performing-Piano"
vel_dyn$Group[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Forte"] <- "Teaching-Forte"
vel_dyn$Group[vel_dyn$Condition == "teaching" & vel_dyn$Subcomponent == "Piano"] <- "Teaching-Piano"
# pairwise comparison
vel_dyn_pairwise <- pairwise.t.test(vel_dyn$Mean, vel_dyn$Group, paired = TRUE, p.adjust.method = "BH")
vel_dyn_pairwise$p.value <- round(vel_dyn_pairwise$p.value, 4)
print(vel_dyn_pairwise)
```

## 3. Trial by trial - individual{#vel_dyn_ind_trial}
- Average KV for each trial for each participant.

```{r vel_dyn_trial, echo = FALSE}
# For each individual
vel_dyn_trial <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*TrialNr, data = df_vel_dyn,
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_dyn_trial <- cbind(vel_dyn_trial[,1:5], as.data.frame(vel_dyn_trial[,6]))
# Change colnames
colnames(vel_dyn_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(vel_dyn_trial)
```

### Plot
#### Line graph
```{r vel_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
p_vel_dyn_trial_forte <- ggplot(data = subset(vel_dyn_trial, vel_dyn_trial$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_dyn_trial_forte
```

```{r vel_dyn_trial_line_piano, echo = FALSE, fig.height = 4}
p_vel_dyn_trial_piano <- ggplot(data = subset(vel_dyn_trial, vel_dyn_trial$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_dyn_trial_piano
```

## 4. Trial by trial - group{#vel_dyn_group_trial}
- Average KV for each trial for each condition.

```{r vel_dyn_trial_all, echo = FALSE}
# Group mean
vel_dyn_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = vel_dyn_trial,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_dyn_trial_all <- cbind(vel_dyn_trial_all[,1:4], as.data.frame(vel_dyn_trial_all[,5]))
# Change colnames
colnames(vel_dyn_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
print(vel_dyn_trial_all)
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r vel_dyn_trial_all_line_forte, echo = FALSE}
p_vel_dyn_trial_all_forte <- ggplot(data = subset(vel_dyn_trial_all, vel_dyn_trial_all$Subcomponent == "Forte"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_dyn_trial_all_forte
```

```{r vel_dyn_trial_all_line_piano, echo = FALSE}
p_vel_dyn_trial_all_piano <- ggplot(data = subset(vel_dyn_trial_all, vel_dyn_trial_all$Subcomponent == "Piano"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_dyn_trial_all_piano
```

# Key Velocity Difference (Difference) - Dynamics
## 1. All - individual{#vel_acc_dyn_ind_all}
- Average KV-Diff for each participant.

```{r vel_acc_dyn, echo = FALSE}
# For each individual
vel_acc_dyn <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_acc_dyn <- cbind(vel_acc_dyn[,1:4], as.data.frame(vel_acc_dyn[,5]))
# Change colnames
colnames(vel_acc_dyn) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(vel_acc_dyn)
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r vel_acc_dyn_bar,  echo = FALSE}
p_vel_acc_dyn <- ggplot(data = vel_acc_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
p_vel_acc_dyn
```

## 2. All - group{#vel_acc_dyn_group_all}
- Average KV-Diff for each condition.

```{r vel_acc_dyn_all, echo = FALSE}
# Group mean
vel_acc_dyn_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = subset(vel_acc_dyn, vel_acc_dyn$SubNr != 3 & vel_acc_dyn$SubNr != 14), # exclude SubNr 3 & 14 as one of the conditions is completely missing
                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_acc_dyn_all <- cbind(vel_acc_dyn_all[,1:3], as.data.frame(vel_acc_dyn_all[,4]))
# Change colnames
colnames(vel_acc_dyn_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
print(vel_acc_dyn_all)
```

### Plot
#### Bar chart
```{r vel_acc_dyn_all_bar,  echo = FALSE}
p_vel_acc_dyn_all <- ggplot(data = vel_acc_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
p_vel_acc_dyn_all
```

#### Box plot
```{r vel_acc_dyn_all_box, echo = FALSE}
p_vel_acc_dyn_all_box <- ggplot(data = subset(vel_acc_dyn, vel_acc_dyn$SubNr != 3 & vel_acc_dyn$SubNr != 14), aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "Difference", subtitle = "KV-Diff: Dynamics")
p_vel_acc_dyn_all_box
```

### Statistics
#### Normality check
```{r normality2, echo = FALSE}
# exclude SubNr 3 & 14
vel_acc_dyn <- subset(vel_acc_dyn, vel_acc_dyn$SubNr != 3 & vel_acc_dyn$SubNr != 14)
# Reduce unused factors
vel_acc_dyn$Subcomponent <- factor(vel_acc_dyn$Subcomponent)
vel_acc_dyn_norm <- by(vel_acc_dyn$Mean, list(vel_acc_dyn$Condition, vel_acc_dyn$Subcomponent), shapiro.test)
print(vel_acc_dyn_norm)
```

#### Two-way ANOVA (ez)
```{r stats3, echo = FALSE}
df_vel_acc_dyn_exc <- subset(df_vel_acc_dyn, df_vel_acc_dyn$SubNr != 3 & df_vel_acc_dyn$SubNr != 14)
vel_acc_dyn_aov <- ezANOVA(
  data = subset(df_vel_acc_dyn_exc, df_vel_acc_dyn_exc$Subcomponent == "FtoP" | df_vel_acc_dyn_exc$Subcomponent == "PtoF")
  , dv = .(Acc)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_acc_dyn_aov$ANOVA$F <- round(vel_acc_dyn_aov$ANOVA$F, 4)
vel_acc_dyn_aov$ANOVA$p <- round(vel_acc_dyn_aov$ANOVA$p, 4)
print(vel_acc_dyn_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats4, echo = FALSE}
vel_acc_dyn_aov_2 <- with(vel_acc_dyn,
                      aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent))))
print(summary(vel_acc_dyn_aov_2))
```

#### Post-hoc (direct comparison)
```{r post-hoc3, echo = FALSE}
vel_acc_ttest_ftop <- t.test(vel_acc_dyn$Mean[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "FtoP"], vel_acc_dyn$Mean[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "FtoP"], paired = TRUE)
vel_acc_ttest_ptof <- t.test(vel_acc_dyn$Mean[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "PtoF"], vel_acc_dyn$Mean[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "PtoF"], paired = TRUE)
# Effect size
vel_acc_ttest_ftop_cohend <- cohen.d(vel_acc_dyn$Mean[vel_acc_dyn$Subcomponent == "FtoP"], vel_acc_dyn$Condition[vel_acc_dyn$Subcomponent == "FtoP"], paired = TRUE)
vel_acc_ttest_ptof_cohend <- cohen.d(vel_acc_dyn$Mean[vel_acc_dyn$Subcomponent == "PtoF"], vel_acc_dyn$Condition[vel_acc_dyn$Subcomponent == "PtoF"], paired = TRUE)

print(vel_acc_ttest_ftop)
print(vel_acc_ttest_ptof)
print(vel_acc_ttest_ftop_cohend)
print(vel_acc_ttest_ptof_cohend)
```

#### Post-hoc (pairwise t test)
```{r post-hoc4, echo = FALSE}
# Create a group label for a post-hoc test
vel_acc_dyn$Group[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "FtoP"] <- "Performing-FtoP" 
vel_acc_dyn$Group[vel_acc_dyn$Condition == "performing" & vel_acc_dyn$Subcomponent == "PtoF"] <- "Performing-PtoF"
vel_acc_dyn$Group[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "FtoP"] <- "Teaching-FtoP"
vel_acc_dyn$Group[vel_acc_dyn$Condition == "teaching" & vel_acc_dyn$Subcomponent == "PtoF"] <- "Teaching-PtoF"
# pairwise comparison
vel_acc_dyn_pairwise <- pairwise.t.test(vel_acc_dyn$Mean, vel_acc_dyn$Group, paired = TRUE, p.adjust.method = "BH")
vel_acc_dyn_pairwise$p.value <- round(vel_acc_dyn_pairwise$p.value, 4)
print(vel_acc_dyn_pairwise)
```

## 3. Trial by trial - individual{#vel_acc_dyn_ind_trial}
- Average KV-Diff for each trial for each participant.

```{r vel_acc_dyn_trial, echo = FALSE}
# For each individual
vel_acc_dyn_trial <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_acc_dyn_trial <- cbind(vel_acc_dyn_trial[,1:5], as.data.frame(vel_acc_dyn_trial[,6]))
# Change colnames
colnames(vel_acc_dyn_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(vel_acc_dyn_trial)
```

### Plot
#### Line graph
```{r vel_acc_dyn_trial_line_ftop, echo = FALSE, fig.height = 4}
p_vel_acc_dyn_trial_ftop <- ggplot(data = subset(vel_acc_dyn_trial, vel_acc_dyn_trial$Subcomponent == "FtoP"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Forte to Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_dyn_trial_ftop
```

```{r vel_acc_dyn_trial_line_ptof, echo = FALSE, fig.height = 4}
p_vel_acc_dyn_trial_ptof <- ggplot(data = subset(vel_acc_dyn_trial, vel_acc_dyn_trial$Subcomponent == "PtoF"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Piano to Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_dyn_trial_ptof
```

## 4. Trial by trial - group{#vel_acc_dyn_group_trial}
- Average KV-Diff for each trial for each condition.

```{r vel_acc_dyn_trial_all, echo = FALSE}
# Group mean
vel_acc_dyn_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = vel_acc_dyn_trial,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_acc_dyn_trial_all <- cbind(vel_acc_dyn_trial_all[,1:4], as.data.frame(vel_acc_dyn_trial_all[,5]))
# Change colnames
colnames(vel_acc_dyn_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
print(vel_acc_dyn_trial_all)
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r vel_acc_dyn_trial_all_line_ftop, echo = FALSE}
p_vel_acc_dyn_trial_all_ftop <- ggplot(data = subset(vel_acc_dyn_trial_all, vel_acc_dyn_trial_all$Subcomponent == "FtoP"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Forte to Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_dyn_trial_all_ftop
```

```{r vel_acc_dyn_trial_all_line_ptof, echo = FALSE}
p_vel_acc_dyn_trial_all_ptof <- ggplot(data = subset(vel_acc_dyn_trial_all, vel_acc_dyn_trial_all$Subcomponent == "PtoF"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Dynamics/Piano to Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_dyn_trial_all_ptof
```

# Key Velocity (KV) - Articulation
## 1. All - individual{#vel_art_ind_all}
- Average KV for each participant.

```{r vel_art, echo = FALSE}
# For each individual
vel_art <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent, data = df_vel_art,
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_art <- cbind(vel_art[,1:4], as.data.frame(vel_art[,5]))
# Change colnames
colnames(vel_art) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(vel_art)
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r vel_art_bar,  echo = FALSE}
p_vel_art <- ggplot(data = vel_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KV (0-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100))
p_vel_art
```

## 2. All - group{#vel_art_group_all}
- Average KV for each condition.

```{r vel_art_all, echo = FALSE}
# Group mean
vel_art_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = vel_art,
                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_art_all <- cbind(vel_art_all[,1:3], as.data.frame(vel_art_all[,4]))
# Change colnames
colnames(vel_art_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
print(vel_art_all)
```

### Plot
#### Bar chart
- Error bars show SEM of the mean

```{r vel_art_all_bar, echo = FALSE}　
p_vel_art_all <- ggplot(data = vel_art_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KV (0-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100))
p_vel_art_all
```

#### Box plot
```{r vel_art_all_box, echo = FALSE}
p_vel_art_all_box <- ggplot(data = vel_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KV (0-127)", subtitle = "KV: Articulation")
p_vel_art_all_box
```

### Statistics
#### Normality check
```{r normality3, echo = FALSE}
# Reduce unused factors
vel_art$Subcomponent <- factor(vel_art$Subcomponent)
vel_art_norm <- by(vel_art$Mean, list(vel_art$Condition, vel_art$Subcomponent), shapiro.test)
print(vel_art_norm)
```

#### Two-way ANOVA (ez)
```{r stats5, echo = FALSE}
vel_art_aov <- ezANOVA(
  data = df_vel_art
  , dv = .(Velocity)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_art_aov$ANOVA$F <- round(vel_art_aov$ANOVA$F, 4)
vel_art_aov$ANOVA$p <- round(vel_art_aov$ANOVA$p, 4)
print(vel_art_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats6, echo = FALSE}
vel_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_art)
print(summary(vel_art_aov_2))
```

#### Post-hoc (direct comparison)
```{r post-hoc5, echo = FALSE}
vel_ttest_legato <- t.test(vel_art$Mean[vel_art$Condition == "performing" & vel_art$Subcomponent == "Legato"], vel_art$Mean[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Legato"], paired = TRUE)
vel_ttest_staccato <- t.test(vel_art$Mean[vel_art$Condition == "performing" & vel_art$Subcomponent == "Staccato"], vel_art$Mean[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Staccato"], paired = TRUE)
# Effect size
vel_ttest_legato_cohend <- cohen.d(vel_art$Mean[vel_art$Subcomponent == "Legato"], vel_art$Condition[vel_art$Subcomponent == "Legato"], paired = TRUE)
vel_ttest_staccato_cohend <- cohen.d(vel_art$Mean[vel_art$Subcomponent == "Staccato"], vel_art$Condition[vel_art$Subcomponent == "Staccato"], paired = TRUE)

print(vel_ttest_legato)
print(vel_ttest_staccato)
print(vel_ttest_legato_cohend)
print(vel_ttest_staccato_cohend)
```

#### Post-hoc (pairwise t test)
```{r post-hoc6, echo = FALSE}
# Create a group label for a post-hoc test
vel_art$Group[vel_art$Condition == "performing" & vel_art$Subcomponent == "Legato"] <- "Performing-Legato" 
vel_art$Group[vel_art$Condition == "performing" & vel_art$Subcomponent == "Staccato"] <- "Performing-Staccato"
vel_art$Group[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Legato"] <- "Teaching-Legato"
vel_art$Group[vel_art$Condition == "teaching" & vel_art$Subcomponent == "Staccato"] <- "Teaching-Staccato"
# pairwise comparison
vel_art_pairwise <- pairwise.t.test(vel_art$Mean, vel_art$Group, paired = TRUE, p.adjust.method = "BH")
vel_art_pairwise$p.value <- round(vel_art_pairwise$p.value, 4)
print(vel_art_pairwise)
```

## 3. Trial by trial - individual{#vel_art_ind_trial}
- Average KV for each trial for each participant.

```{r vel_art_trial, echo = FALSE}
# For each individual
vel_art_trial <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*TrialNr, data = df_vel_art,
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_art_trial <- cbind(vel_art_trial[,1:5], as.data.frame(vel_art_trial[,6]))
# Change colnames
colnames(vel_art_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(vel_art_trial)
```

### Plot
#### Line graph
```{r vel_art_trial_line_legato, echo = FALSE, fig.height = 4}
p_vel_art_trial_legato <- ggplot(data = subset(vel_art_trial, vel_art_trial$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_art_trial_legato
```

```{r vel_art_trial_line_piano, echo = FALSE, fig.height = 4}
p_vel_art_trial_staccato <- ggplot(data = subset(vel_art_trial, vel_art_trial$Subcomponent == "Staccato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_art_trial_staccato
```

## 4. Trial by trial - group{#vel_art_group_trial}
- Average KV for each trial for each condition.

```{r vel_art_trial_all, echo = FALSE}
# Group mean
vel_art_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = vel_art_trial,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_art_trial_all <- cbind(vel_art_trial_all[,1:4], as.data.frame(vel_art_trial_all[,5]))
# Change colnames
colnames(vel_art_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
print(vel_art_trial_all)
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r vel_art_trial_all_line_legato, echo = FALSE}
p_vel_art_trial_all_legato <- ggplot(data = subset(vel_art_trial_all, vel_art_trial_all$Subcomponent == "Legato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_art_trial_all_legato
```

```{r vel_art_trial_all_line_staccato, echo = FALSE}
p_vel_art_trial_all_staccato <- ggplot(data = subset(vel_art_trial_all, vel_art_trial_all$Subcomponent == "Staccato"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KV (0-127)", subtitle = "KV: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_art_trial_all_staccato
```

# Key Velocity Difference (Difference) - Articulation
## 1. All - individual{#vel_acc_art_ind_all}
- Average KV-Diff for each participant.

```{r vel_acc_art, echo = FALSE}
# For each individual
vel_acc_art <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_acc_art <- cbind(vel_acc_art[,1:4], as.data.frame(vel_acc_art[,5]))
# Change colnames
colnames(vel_acc_art) <- c("SubNr", "Condition", "Skill", "Subcomponent", "N", "Mean", "SD")
print(vel_acc_art)
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r vel_acc_art_bar,  echo = FALSE}
p_vel_acc_art <- ggplot(data = vel_acc_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "Difference", subtitle = "KV-Diff: Articulaion")
p_vel_acc_art
```

## 2. All - group{#vel_acc_art_group_all}
- Average KV-Diff for each condition.

```{r vel_acc_art_all, echo = FALSE}
# Group mean
vel_acc_art_all <- aggregate(Mean~Condition*Skill*Subcomponent, data = vel_acc_art,
                         FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_acc_art_all <- cbind(vel_acc_art_all[,1:3], as.data.frame(vel_acc_art_all[,4]))
# Change colnames
colnames(vel_acc_art_all) <- c("Condition", "Skill", "Subcomponent", "N", "Mean", "SD", "SEM")
print(vel_acc_art_all)
```

### Plot
#### Bar chart
```{r vel_acc_art_all_bar,  echo = FALSE}
p_vel_acc_art_all <- ggplot(data = vel_acc_art_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "Difference", subtitle = "KV-Diff: Articulation")
p_vel_acc_art_all
```

#### Box plot
```{r vel_acc_art_all_box, echo = FALSE}
p_vel_acc_art_all_box <- ggplot(data = vel_acc_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "Difference", subtitle = "KV-Diff: Articulation")
p_vel_acc_art_all_box
```

### Statistics
#### Normality check
```{r normality4, echo = FALSE}
# Reduce unused factors
vel_acc_art$Subcomponent <- factor(vel_acc_art$Subcomponent)
vel_acc_art_norm <- by(vel_acc_art$Mean, list(vel_acc_art$Condition, vel_acc_art$Subcomponent), shapiro.test)
print(vel_acc_art_norm)
```

#### Two-way ANOVA (ez)
```{r stats7, echo = FALSE}
vel_acc_art_aov <- ezANOVA(
  data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL")
  , dv = .(Acc)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
vel_acc_art_aov$ANOVA$F <- round(vel_acc_art_aov$ANOVA$F, 4)
vel_acc_art_aov$ANOVA$p <- round(vel_acc_art_aov$ANOVA$p, 4)
print(vel_acc_art_aov$ANOVA)
```

#### Post-hoc (direct comparison)
```{r post-hoc7, echo = FALSE}
vel_acc_ttest_ltos <- t.test(vel_acc_art$Mean[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "LtoS"], vel_acc_art$Mean[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "LtoS"], paired = TRUE)
vel_acc_ttest_stol <- t.test(vel_acc_art$Mean[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "StoL"], vel_acc_art$Mean[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "StoL"], paired = TRUE)
# Effect size
vel_acc_ttest_ltos_cohend <- cohen.d(vel_acc_art$Mean[vel_acc_art$Subcomponent == "LtoS"], vel_acc_art$Condition[vel_acc_art$Subcomponent == "LtoS"], paired = TRUE)
vel_acc_ttest_stol_cohend <- cohen.d(vel_acc_art$Mean[vel_acc_art$Subcomponent == "StoL"], vel_acc_art$Condition[vel_acc_art$Subcomponent == "StoL"], paired = TRUE)

print(vel_acc_ttest_ltos)
print(vel_acc_ttest_stol)
print(vel_acc_ttest_ltos_cohend)
print(vel_acc_ttest_stol_cohend)
```

#### Post-hoc (pairwise t test)
```{r post-hoc8, echo = FALSE}
# Create a group label for a post-hoc test
vel_acc_art$Group[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "LtoS"] <- "Performing-LtoS" 
vel_acc_art$Group[vel_acc_art$Condition == "performing" & vel_acc_art$Subcomponent == "StoL"] <- "Performing-StoL"
vel_acc_art$Group[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "LtoS"] <- "Teaching-LtoS"
vel_acc_art$Group[vel_acc_art$Condition == "teaching" & vel_acc_art$Subcomponent == "StoL"] <- "Teaching-StoL"
# pairwise comparison
vel_acc_art_pairwise <- pairwise.t.test(vel_acc_art$Mean, vel_acc_art$Group, paired = TRUE, p.adjust.method = "BH")
vel_acc_art_pairwise$p.value <- round(vel_acc_art_pairwise$p.value, 4)
print(vel_acc_art_pairwise)
```

#### Two-way ANOVA (aov)
```{r stats8, echo = FALSE}
vel_acc_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_acc_art)
print(summary(vel_acc_art_aov_2))
```

## 3. Trial by trial - individual{#vel_acc_art_ind_trial}
- Average KV-Diff for each trial for each participant.

```{r vel_acc_art_trial, echo = FALSE}
# For each individual
vel_acc_art_trial <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*TrialNr, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"),
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_acc_art_trial <- cbind(vel_acc_art_trial[,1:5], as.data.frame(vel_acc_art_trial[,6]))
# Change colnames
colnames(vel_acc_art_trial) <- c("SubNr", "Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD")
print(vel_acc_art_trial)
```

### Plot
#### Line graph
```{r vel_acc_art_trial_line_ltos, echo = FALSE, fig.height = 4}
p_vel_acc_art_trial_ltos <- ggplot(data = subset(vel_acc_art_trial, vel_acc_art_trial$Subcomponent == "LtoS"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_art_trial_ltos
```

```{r vel_acc_art_trial_line_stol, echo = FALSE, fig.height = 4}
p_vel_acc_art_trial_stol <- ggplot(data = subset(vel_acc_art_trial, vel_acc_art_trial$Subcomponent == "StoL"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Staccato to Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_art_trial_stol
```

## 4. Trial by trial - group{#vel_acc_art_group_trial}
- Average KV-Diff for each trial for each condition.

```{r vel_acc_art_trial_all, echo = FALSE}
# Group mean
vel_acc_art_trial_all <- aggregate(Mean~Condition*Skill*Subcomponent*TrialNr, data = vel_acc_art_trial,
                       FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_acc_art_trial_all <- cbind(vel_acc_art_trial_all[,1:4], as.data.frame(vel_acc_art_trial_all[,5]))
# Change colnames
colnames(vel_acc_art_trial_all) <- c("Condition", "Skill", "Subcomponent", "TrialNr", "N", "Mean", "SD", "SEM")
print(vel_acc_art_trial_all)
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r vel_acc_art_trial_all_line_ltos, echo = FALSE}
p_vel_acc_art_trial_all_ltos <- ggplot(data = subset(vel_acc_art_trial_all, vel_acc_art_trial_all$Subcomponent == "LtoS"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_art_trial_all_ltos
```

```{r vel_acc_art_trial_all_line_stol, echo = FALSE}
p_vel_acc_art_trial_all_stol <- ggplot(data = subset(vel_acc_art_trial_all, vel_acc_art_trial_all$Subcomponent == "StoL"), aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "Difference", subtitle = "KV-Diff: Articulation/Legato to Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_vel_acc_art_trial_all_stol
```

<!-- # KV-Order-Dynamics -->
<!-- - Investigate whether there is any order effect of the conditions. -->

<!-- ## 1. All{#vel_dyn_order_group_all} -->
<!-- ```{r vel_dyn_order_all, echo = all} -->
<!-- # For each participant -->
<!-- vel_dyn_order <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*Order, data = df_vel_dyn, -->
<!--                                FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
<!-- vel_dyn_order <- cbind(vel_dyn_order[,1:5], as.data.frame(vel_dyn_order[,6])) -->
<!-- # Change colnames -->
<!-- colnames(vel_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->

<!-- # Group mean -->
<!-- vel_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_dyn_order, -->
<!--                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
<!-- vel_dyn_order_all <- cbind(vel_dyn_order_all[,1:4], as.data.frame(vel_dyn_order_all[,5])) -->
<!-- # Change colnames -->
<!-- colnames(vel_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
<!-- print(vel_dyn_order_all) -->
<!-- ``` -->

<!-- ### Plot -->
<!-- #### Bar graph -->
<!-- - Error bars show SEM of the mean -->

<!-- ```{r vel_dyn_order_all_bar, echo = FALSE} -->
<!-- p_vel_dyn_order_all <- ggplot(data = vel_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
<!--                 width = .2, position = position_dodge(.9)) + -->
<!--   facet_wrap(Order ~ .) + -->
<!--   labs(y = "KV (0-127)", subtitle = "KV: Dynamics") + coord_cartesian(ylim = c(40, 100)) -->
<!-- p_vel_dyn_order_all -->
<!-- ``` -->

<!-- #### Box plot -->
<!-- ```{r vel_dyn_order_all_box, echo = FALSE} -->
<!-- p_vel_dyn_order_all_box <- ggplot(data = vel_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
<!--   geom_boxplot() + -->
<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
<!--   facet_wrap(Order ~ .) + -->
<!--   labs(y = "KV (0-127)", subtitle = "KV: Dynamics")  -->
<!-- p_vel_dyn_order_all_box -->
<!-- ``` -->

<!-- ### Statistics -->
<!-- #### Normality check -->
<!-- ```{r normality5, echo = FALSE} -->
<!-- # Normality check -->
<!-- vel_dyn_order$Subcomponent <- factor(vel_dyn_order$Subcomponent) -->
<!-- vel_dyn_order_norm <- by(vel_dyn_order$Mean, list(vel_dyn_order$Condition, vel_dyn_order$Subcomponent, vel_dyn_order$Order), shapiro.test) -->
<!-- print(vel_dyn_order_norm) -->
<!-- ``` -->

<!-- #### Three-way ANOVA (ez) -->
<!-- ```{r stats9, echo = FALSE} -->
<!-- vel_dyn_order_aov <- ezANOVA( -->
<!--   data = df_vel_dyn -->
<!--   , dv = .(Velocity) -->
<!--   , wid = .(SubNr) -->
<!--   , between = .(Order) -->
<!--   , within = .(Condition, Subcomponent) -->
<!--   , type = 3 -->
<!--   , detailed = TRUE -->
<!-- ) -->
<!-- vel_dyn_order_aov$ANOVA$F <- round(vel_dyn_order_aov$ANOVA$F, 4) -->
<!-- vel_dyn_order_aov$ANOVA$p <- round(vel_dyn_order_aov$ANOVA$p, 4) -->
<!-- print(vel_dyn_order_aov) -->
<!-- ``` -->

<!-- # KV-Diff-Order-Dynamics -->
<!-- - Investigate whether there is any order effect of the conditions. -->

<!-- ## 1. All{#vel_acc_dyn_order_group_all} -->
<!-- ```{r vel_acc_dyn_order_all, echo = all} -->
<!-- # For each participant -->
<!-- vel_acc_dyn_order <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF"), -->
<!--                                FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
<!-- vel_acc_dyn_order <- cbind(vel_acc_dyn_order[,1:5], as.data.frame(vel_acc_dyn_order[,6])) -->
<!-- # Change colnames -->
<!-- colnames(vel_acc_dyn_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->

<!-- # Group mean -->
<!-- vel_acc_dyn_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_acc_dyn_order, -->
<!--                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
<!-- vel_acc_dyn_order_all <- cbind(vel_acc_dyn_order_all[,1:4], as.data.frame(vel_acc_dyn_order_all[,5])) -->
<!-- # Change colnames -->
<!-- colnames(vel_acc_dyn_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
<!-- print(vel_acc_dyn_order_all) -->
<!-- ``` -->

<!-- ### Plot -->
<!-- #### Bar graph -->
<!-- - Error bars show SEM of the mean -->

<!-- ```{r vel_acc_dyn_order_all_bar, echo = FALSE} -->
<!-- p_vel_acc_dyn_order_all <- ggplot(data = vel_acc_dyn_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
<!--                 width = .2, position = position_dodge(.9)) + -->
<!--   facet_wrap(Order ~ .) + -->
<!--   labs(y = "Difference", subtitle = "KV-Diff: Dynamics") -->
<!-- p_vel_acc_dyn_order_all -->
<!-- ``` -->

<!-- #### Box plot -->
<!-- ```{r vel_acc_dyn_order_all_box, echo = FALSE} -->
<!-- p_vel_acc_dyn_order_all_box <- ggplot(data = vel_acc_dyn_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
<!--   geom_boxplot() + -->
<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
<!--   facet_wrap(Order ~ .) + -->
<!--   labs(y = "Difference", subtitle = "KV-Diff: Dynamics")  -->
<!-- p_vel_acc_dyn_order_all_box -->
<!-- ``` -->

<!-- ### Statistics -->
<!-- #### Normality check -->
<!-- ```{r normality6, echo = FALSE} -->
<!-- # Normality check -->
<!-- vel_acc_dyn_order$Subcomponent <- factor(vel_acc_dyn_order$Subcomponent) -->
<!-- vel_acc_dyn_order_norm <- by(vel_acc_dyn_order$Mean, list(vel_acc_dyn_order$Condition, vel_acc_dyn_order$Subcomponent, vel_acc_dyn_order$Order), shapiro.test) -->
<!-- print(vel_acc_dyn_order_norm) -->
<!-- ``` -->

<!-- #### Three-way ANOVA (ez) -->
<!-- ```{r stats10, echo = FALSE} -->
<!-- vel_acc_dyn_order_aov <- ezANOVA( -->
<!--   data = subset(df_vel_acc_dyn, df_vel_acc_dyn$Subcomponent == "FtoP" | df_vel_acc_dyn$Subcomponent == "PtoF") -->
<!--   , dv = .(Acc) -->
<!--   , wid = .(SubNr) -->
<!--   , between = .(Order) -->
<!--   , within = .(Condition, Subcomponent) -->
<!--   , type = 3 -->
<!--   , detailed = TRUE -->
<!-- ) -->
<!-- vel_acc_dyn_order_aov$ANOVA$F <- round(vel_acc_dyn_order_aov$ANOVA$F, 4) -->
<!-- vel_acc_dyn_order_aov$ANOVA$p <- round(vel_acc_dyn_order_aov$ANOVA$p, 4) -->
<!-- print(vel_acc_dyn_order_aov) -->
<!-- ``` -->

<!-- # KV-Order-Articulation -->
<!-- - Investigate whether there is any order effect of the conditions. -->

<!-- ## 1. All{#vel_art_order_group_all} -->
<!-- ```{r vel_art_order_all, echo = all} -->
<!-- # For each participant -->
<!-- vel_art_order <- aggregate(Velocity~SubNr*Condition*Skill*Subcomponent*Order, data = df_vel_art, -->
<!--                                FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
<!-- vel_art_order <- cbind(vel_art_order[,1:5], as.data.frame(vel_art_order[,6])) -->
<!-- # Change colnames -->
<!-- colnames(vel_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->

<!-- # Group mean -->
<!-- vel_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_art_order, -->
<!--                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
<!-- vel_art_order_all <- cbind(vel_art_order_all[,1:4], as.data.frame(vel_art_order_all[,5])) -->
<!-- # Change colnames -->
<!-- colnames(vel_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
<!-- print(vel_art_order_all) -->
<!-- ``` -->

<!-- ### Plot -->
<!-- #### Bar graph -->
<!-- - Error bars show SEM of the mean -->

<!-- ```{r vel_art_order_all_bar, echo = FALSE} -->
<!-- p_vel_art_order_all <- ggplot(data = vel_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
<!--                 width = .2, position = position_dodge(.9)) + -->
<!--   facet_wrap(Order ~ .) + -->
<!--   labs(y = "KV (0-127)", subtitle = "KV: Articulation") + coord_cartesian(ylim = c(40, 100)) -->
<!-- p_vel_art_order_all -->
<!-- ``` -->

<!-- #### Box plot -->
<!-- ```{r vel_art_order_all_box, echo = FALSE} -->
<!-- p_vel_art_order_all_box <- ggplot(data = vel_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
<!--   geom_boxplot() + -->
<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
<!--   facet_wrap(Order ~ .) + -->
<!--   labs(y = "KV (0-127)", subtitle = "KV: Articulation")  -->
<!-- p_vel_art_order_all_box -->
<!-- ``` -->

<!-- ### Statistics -->
<!-- #### Normality check -->
<!-- ```{r normality7, echo = FALSE} -->
<!-- # Normality check -->
<!-- vel_art_order$Subcomponent <- factor(vel_art_order$Subcomponent) -->
<!-- vel_art_order_norm <- by(vel_art_order$Mean, list(vel_art_order$Condition, vel_art_order$Subcomponent, vel_art_order$Order), shapiro.test) -->
<!-- print(vel_art_order_norm) -->
<!-- ``` -->

<!-- #### Three-way ANOVA (ez) -->
<!-- ```{r stats11, echo = FALSE} -->
<!-- vel_art_order_aov <- ezANOVA( -->
<!--   data = df_vel_art -->
<!--   , dv = .(Velocity) -->
<!--   , wid = .(SubNr) -->
<!--   , between = .(Order) -->
<!--   , within = .(Condition, Subcomponent) -->
<!--   , type = 3 -->
<!--   , detailed = TRUE -->
<!-- ) -->
<!-- vel_art_order_aov$ANOVA$F <- round(vel_art_order_aov$ANOVA$F, 4) -->
<!-- vel_art_order_aov$ANOVA$p <- round(vel_art_order_aov$ANOVA$p, 4) -->
<!-- print(vel_art_order_aov) -->
<!-- ``` -->

<!-- # KV-Diff-Order-Articulation -->
<!-- - Investigate whether there is any order effect of the conditions. -->

<!-- ## 1. All{#vel_acc_art_order_group_all} -->
<!-- ```{r vel_acc_art_order_all, echo = all} -->
<!-- # For each participant -->
<!-- vel_acc_art_order <- aggregate(Acc~SubNr*Condition*Skill*Subcomponent*Order, data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL"), -->
<!--                                FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))}) -->
<!-- vel_acc_art_order <- cbind(vel_acc_art_order[,1:5], as.data.frame(vel_acc_art_order[,6])) -->
<!-- # Change colnames -->
<!-- colnames(vel_acc_art_order) <- c("SubNr", "Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD") -->

<!-- # Group mean -->
<!-- vel_acc_art_order_all <- aggregate(Mean~Condition*Skill*Subcomponent*Order, data = vel_acc_art_order, -->
<!--                          FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)}) -->
<!-- vel_acc_art_order_all <- cbind(vel_acc_art_order_all[,1:4], as.data.frame(vel_acc_art_order_all[,5])) -->
<!-- # Change colnames -->
<!-- colnames(vel_acc_art_order_all) <- c("Condition", "Skill", "Subcomponent", "Order", "N", "Mean", "SD", "SEM") -->
<!-- print(vel_acc_art_order_all) -->
<!-- ``` -->

<!-- ### Plot -->
<!-- #### Bar graph -->
<!-- - Error bars show SEM of the mean -->

<!-- ```{r vel_acc_art_order_all_bar, echo = FALSE} -->
<!-- p_vel_acc_art_order_all <- ggplot(data = vel_acc_art_order_all, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
<!--   geom_bar(stat = "identity", position = position_dodge()) + -->
<!--   geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), -->
<!--                 width = .2, position = position_dodge(.9)) + -->
<!--   facet_wrap(Order ~ .) + -->
<!--   labs(y = "Difference", subtitle = "KV-Diff: Articulation") -->
<!-- p_vel_acc_art_order_all -->
<!-- ``` -->

<!-- #### Box plot -->
<!-- ```{r vel_acc_art_order_all_box, echo = FALSE} -->
<!-- p_vel_acc_art_order_all_box <- ggplot(data = vel_acc_art_order, aes(x = Subcomponent, y = Mean, fill = Condition)) + -->
<!--   geom_boxplot() + -->
<!--   geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(.75)) + -->
<!--   facet_wrap(Order ~ .) + -->
<!--   labs(y = "Difference", subtitle = "KV: Articulation")  -->
<!-- p_vel_acc_art_order_all_box -->
<!-- ``` -->

<!-- ### Statistics -->
<!-- #### Normality check -->
<!-- ```{r normality8, echo = FALSE} -->
<!-- # Normality check -->
<!-- vel_acc_art_order$Subcomponent <- factor(vel_acc_art_order$Subcomponent) -->
<!-- vel_acc_art_order_norm <- by(vel_acc_art_order$Mean, list(vel_acc_art_order$Condition, vel_acc_art_order$Subcomponent, vel_acc_art_order$Order), shapiro.test) -->
<!-- print(vel_acc_art_order_norm) -->
<!-- ``` -->

<!-- #### Three-way ANOVA (ez) -->
<!-- ```{r stats12, echo = FALSE} -->
<!-- vel_acc_art_order_aov <- ezANOVA( -->
<!--   data = subset(df_vel_acc_art, df_vel_acc_art$Subcomponent == "LtoS" | df_vel_acc_art$Subcomponent == "StoL") -->
<!--   , dv = .(Acc) -->
<!--   , wid = .(SubNr) -->
<!--   , between = .(Order) -->
<!--   , within = .(Condition, Subcomponent) -->
<!--   , type = 3 -->
<!--   , detailed = TRUE -->
<!-- ) -->
<!-- vel_acc_art_order_aov$ANOVA$F <- round(vel_acc_art_order_aov$ANOVA$F, 4) -->
<!-- vel_acc_art_order_aov$ANOVA$p <- round(vel_acc_art_order_aov$ANOVA$p, 4) -->
<!-- print(vel_acc_art_order_aov) -->
<!-- ``` -->

# Sequence plot
### Dynamics{#sequence_dyn}
```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE}
# Average vel_dynocity for each note
# For each participant
vel_dyn_seq <- aggregate(Velocity~SubNr*Condition*Skill*Note, data = df_vel_dyn, 
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_dyn_seq <- cbind(vel_dyn_seq[,1:4], as.data.frame(vel_dyn_seq[,5]))
# Change colnames
colnames(vel_dyn_seq) <- c("SubNr", "Condition", "Skill", "Note", "N", "Mean", "SD")

# Group mean
vel_dyn_seq_stats <- aggregate(Mean~Condition*Skill*Note, data = vel_dyn_seq,
                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_dyn_seq_stats <- cbind(vel_dyn_seq_stats[,1:3], as.data.frame(vel_dyn_seq_stats[,4]))
# Change colnames
colnames(vel_dyn_seq_stats) <- c("Condition", "Skill", "Note", "N", "Mean", "SD", "SEM")

# Average acceleration for each note
# For each participant
vel_dyn_acc_seq <- aggregate(Acc~SubNr*Condition*Skill*Interval, data = df_vel_acc_dyn, 
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_dyn_acc_seq <- cbind(vel_dyn_acc_seq[,1:4], as.data.frame(vel_dyn_acc_seq[,5]))
# Change colnames
colnames(vel_dyn_acc_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")

# Group mean
vel_dyn_acc_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = vel_dyn_acc_seq,
                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_dyn_acc_seq_stats <- cbind(vel_dyn_acc_seq_stats[,1:3], as.data.frame(vel_dyn_acc_seq_stats[,4]))
# Change colnames
colnames(vel_dyn_acc_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")

p_vel_dyn_seq <- ggplot(data = vel_dyn_seq_stats, aes(x = Note, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
                position = position_dodge(0.05)) + 
  labs(y = "Velocity (0-127)") + scale_x_continuous(breaks=seq(1,67,1))
p_vel_dyn_seq

p_vel_dyn_acc_seq <- ggplot(data = vel_dyn_acc_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
                position = position_dodge(.05)) + 
  labs(x = "Interval", y = "Difference") + scale_x_continuous(breaks=seq(1,66,1))
p_vel_dyn_acc_seq
```

### Articulation{#sequence_art}
```{r seq_art, fig.width = 10, fig.height = 2, echo = FALSE}
# Average vel_artocity for each note
# For each participant
vel_art_seq <- aggregate(Velocity~SubNr*Condition*Skill*Note, data = df_vel_art, 
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_art_seq <- cbind(vel_art_seq[,1:4], as.data.frame(vel_art_seq[,5]))
# Change colnames
colnames(vel_art_seq) <- c("SubNr", "Condition", "Skill", "Note", "N", "Mean", "SD")

# Group mean
vel_art_seq_stats <- aggregate(Mean~Condition*Skill*Note, data = vel_art_seq,
                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_art_seq_stats <- cbind(vel_art_seq_stats[,1:3], as.data.frame(vel_art_seq_stats[,4]))
# Change colnames
colnames(vel_art_seq_stats) <- c("Condition", "Skill", "Note", "N", "Mean", "SD", "SEM")

# Average acceleration for each note
# For each participant
vel_art_acc_seq <- aggregate(Acc~SubNr*Condition*Skill*Interval, data = df_vel_acc_art, 
                     FUN = function(x){c(N = length(x), mean = mean(x), sd = sd(x))})
vel_art_acc_seq <- cbind(vel_art_acc_seq[,1:4], as.data.frame(vel_art_acc_seq[,5]))
# Change colnames
colnames(vel_art_acc_seq) <- c("SubNr", "Condition", "Skill", "Interval", "N", "Mean", "SD")

# Group mean
vel_art_acc_seq_stats <- aggregate(Mean~Condition*Skill*Interval, data = vel_art_acc_seq,
                           FUN = function(x){round(c(N = length(x), mean = mean(x), sd = sd(x), sem = sd(x)/sqrt(length(x))), 4)})
vel_art_acc_seq_stats <- cbind(vel_art_acc_seq_stats[,1:3], as.data.frame(vel_art_acc_seq_stats[,4]))
# Change colnames
colnames(vel_art_acc_seq_stats) <- c("Condition", "Skill", "Interval", "N", "Mean", "SD", "SEM")

p_vel_art_seq <- ggplot(data = vel_art_seq_stats, aes(x = Note, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
                position = position_dodge(0.05)) + 
  labs(y = "Velocity (0-127)") + scale_x_continuous(breaks=seq(1,67,1))
p_vel_art_seq

p_vel_art_acc_seq <- ggplot(data = vel_art_acc_seq_stats, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
                position = position_dodge(.05)) + 
  labs(x = "Interval", y = "Difference") + scale_x_continuous(breaks=seq(1,66,1))
p_vel_art_acc_seq
```