### 2020-11-18 script/RNotebook_2/ioi.Rmd
stats for ioi
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Tempo Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of all analyses for a tempo.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Outline - Interonset intervals (IOIs)
## [Experimental Design](#design)

## [Articulation](#ioi-articulation)

- Individual results
  + [All](#art_ind_all)
  + [Phrase bondaries](#art_ind_phrase)
  + [Trial by trial](#art_ind_trial)

- Group results and stats (if available)
  + [All](#art_group_all)
  + [Phrase bondaries](#art_group_phrase)
  + [Trial by trial](#art_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#art_order_group_all)

- [Sequence plot](#sequence_art)

## [Dynamics](#ioi-dynamics)

- Individual results
  + [All](#dyn_ind_all)
  + [Phrase bondaries](#dyn_ind_phrase)
  + [Trial by trial](#dyn_ind_trial)

- Group results and stats (if available)
  + [All](#dyn_group_all)
  + [Phrase bondaries](#dyn_group_phrase)
  + [Trial by trial](#dyn_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#dyn_order_group_all)

- [Sequence plot](#sequence_dyn)

## Experimental Design{#design}
- Independent variables: Condition (2 levels: teaching vs. performing) x Skill (2 levels: articulation vs. dynamics)
- Dependent variables: Interonset intervals, Variability (SD/mean IOI)

## Main predictions
Participants will play slower while teaching regardless of the skills.

## Excluded participnats
- SubNr 1: Most of responses (more than 90%) were removed due to outliers
- SubNr 3: Experimental error
- SubNr 8: Experimental error
- SubNr 14: Experimental error

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
# statistics
if (!require("stats")) {install.packages("stats"); require("stats")}
if (!require("effsize")) {install.packages("effsize"); require("effsize")}
if (!require("ez")) {install.packages("ez"); require("ez")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r file, include = FALSE}
filename_ioi = "../R/preprocessor/trimmed/data_ioi_1.txt"
```

```{r extract, include = FALSE}
dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv

# SubNr as a factor
dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)

# Exclude SubNr 1 because most responses were removed because of outliers
dt_ioi <- dt_ioi[SubNr != 1]

# Include only articulation
dt_ioi_art <- dt_ioi[Skill == "articulation"]
# Include only dynamics
dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
```

# IOI-Articulation 
## 1. All - individual{#art_ind_all}
- Average IOIs for each participant.

```{r ioi_art, echo = FALSE}
# For each individual
ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
ioi_art
```

### Plot
#### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_art_bar,  echo = FALSE}
p_ioi_art <- ggplot(data = ioi_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 18, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 280))
p_ioi_art
```

## 2. All - group{#art_group_all}
- Average IOIs for each condition.
- Hypothesis: Participants will play slower when teaching.

```{r ioi_art_all, echo = FALSE}
# Group mean
ioi_art_all <- ioi_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_art)/2)), by = .(Condition, Skill)]
ioi_art_all
```

### Plot
#### Bar chart
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_art_all_bar, echo = FALSE}
p_ioi_art_all <- ggplot(data = ioi_art_all, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(120, 210)) +
  theme(legend.position = "None")
p_ioi_art_all
```

#### Box plot
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_art_all_box, echo = FALSE}
p_ioi_art_all_box <- ggplot(data = ioi_art, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + #coord_cartesian(ylim = c(100, 210)) +
  theme(legend.position = "None")
p_ioi_art_all_box
```

### Statistics
#### Normality check
- Violate the normality
```{r normality1, echo = FALSE}
# Normality check
ioi_art_norm <- by(ioi_art$Mean, list(ioi_art$Condition), shapiro.test)
print(ioi_art_norm)
```

#### Paired t-test
```{r stats1, echo = FALSE}
ioi_art_ttest <- t.test(ioi_art[Condition == "teaching"]$Mean, ioi_art[Condition == "performing"]$Mean, paired = TRUE)
print(ioi_art_ttest)
```

#### Wilcoxon test
```{r stats2, echo = FALSE}
ioi_art_wilcoxon <- wilcox.test(Mean ~ Condition, data = ioi_art, paired = TRUE)
ioi_art_wilcoxon
```

## 3. Phrase boundaries - individual{#art_ind_phrase}
- Compare IOIs at phrase boundaries with others fir each participant.

```{r ioi_art_ch, echo = FALSE}
# Define phrase boundaries
dt_ioi_art$Boundary <- "No"
dt_ioi_art[Interval == 8 | Interval == 16 | Interval == 24 | Interval == 41 | Interval == 49 | Interval == 57]$Boundary <- "Yes"

# For each individual
ioi_art_ch <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Boundary)]
ioi_art_ch
```

### Plot
#### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_art_ch_bar, echo = FALSE}
p_ioi_art_ch <- ggplot(data = ioi_art_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Boundary ~ .) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 18, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 280))
p_ioi_art_ch
```

## 4. Phrase boundaries - group{#art_group_phrase}
- Compare IOIs at phrase boundaries with others in the two conditions.
- Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations.

```{r ioi_art_ch_all, echo = FALSE}
# Group mean
ioi_art_ch_all <- ioi_art_ch[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_art_ch)/2)), by = .(Condition, Skill, Boundary)]
ioi_art_ch_all
```

### Plot
#### Bar chart
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_art_ch_all_bar, echo = FALSE}
p_ioi_art_ch_all_bar <- ggplot(data = ioi_art_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  #facet_grid(. ~ Boundary) +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 210))
p_ioi_art_ch_all_bar
```

#### Box plot
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_art_ch_all_box, echo = FALSE}
p_ioi_art_ch_box <- ggplot(data = ioi_art_ch, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation")
p_ioi_art_ch_box
```

### Statistics
#### Normality check
- Violate the normality
```{r normality2, echo = FALSE}
ioi_art_ch_norm <- by(ioi_art_ch$Mean, list(ioi_art_ch$Condition, ioi_art_ch$Boundary), shapiro.test)
print(ioi_art_ch_norm)
```
#### Two-way ANOVA (ez)
```{r stats3, echo = FALSE}
ioi_art_ch_aov <- ezANOVA(
  data = dt_ioi_art
  , dv = .(IOI)
  , wid = .(SubNr)
  , within = .(Condition, Boundary)
  , type = 3
  , detailed = TRUE
)
ioi_art_ch_aov$ANOVA$F <- round(ioi_art_ch_aov$ANOVA$F, 4)
ioi_art_ch_aov$ANOVA$p <- round(ioi_art_ch_aov$ANOVA$p, 4)
print(ioi_art_ch_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats4, echo = FALSE}
ioi_art_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_art_ch)
print(summary(ioi_art_ch_aov_2))
```

## 5. Trial by trial - individual{#art_ind_trial}
- Average IOIs for each trial for each participant.

```{r ioi_art_trial, echo = FALSE}
# For each individual
ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, TrialNr)]
ioi_art_trial
```

### Plot
#### Line graph
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_art_trial_line, echo = FALSE, fig.height = 4}
p_ioi_art_trial <- ggplot(data = ioi_art_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_art_trial
```

## 6. Trial by trial - group{#art_group_trial}
- Average IOIs for each trial for each condition.

```{r ioi_art_trial_all, echo = FALSE}
# Group mean
ioi_art_trial_all <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_art_trial)/2)), by = .(Condition, Skill, TrialNr)]
ioi_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_art_trial_all_line, echo = FALSE}
p_ioi_art_trial_all <- ggplot(data = ioi_art_trial_all, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_art_trial_all
```

# IOI-Dynamics
## 1. All - individual{#dyn_ind_all}
- Average IOIs for each participant.

```{r ioi_dyn, echo = FALSE}
# For each individual
ioi_dyn <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
ioi_dyn
```

### Plot
#### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_dyn_bar,  echo = FALSE}
p_ioi_dyn <- ggplot(data = ioi_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 18, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 280))
p_ioi_dyn
```

## 2. All - group{#dyn_group_all}
- Average IOIs for each condition.
- Hypothesis: Participants will play slower when teaching.

```{r ioi_dyn_all, echo = FALSE}
# Group mean
ioi_dyn_all <- ioi_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_dyn)/2)), by = .(Condition, Skill)]
ioi_dyn_all
```

### Plot
#### Bar chart
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_dyn_all_bar, echo = FALSE}
p_ioi_dyn_all <- ggplot(data = ioi_dyn_all, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(120, 210)) +
  theme(legend.position = "None")
p_ioi_dyn_all
```

#### Box plot
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_dyn_all_box, echo = FALSE}
p_ioi_dyn_all_box <- ggplot(data = ioi_dyn, aes(x = Condition, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + #coord_cartesian(ylim = c(100, 210)) +
  theme(legend.position = "None")
p_ioi_dyn_all_box
```

### Statistics
#### Normality check
- Violate the normality
```{r normality3, echo = FALSE}
# Normality check
ioi_dyn_norm <- by(ioi_dyn$Mean, list(ioi_dyn$Condition), shapiro.test)
print(ioi_dyn_norm)
```

#### Paired t-test
```{r stats5, echo = FALSE}
ioi_dyn_ttest <- t.test(ioi_dyn$Mean[ioi_dyn$Condition == "teaching"], ioi_dyn$Mean[ioi_dyn$Condition == "performing"], paired = TRUE)
print(ioi_dyn_ttest)
```

#### Wilcoxon test
```{r stats6, echo = FALSE}
ioi_dyn_wilcoxon <- wilcox.test(Mean ~ Condition, data = ioi_dyn, paired = TRUE)
ioi_dyn_wilcoxon
```

## 3. Phrase boundaries - individual{#dyn_ind_phrase}
- Compare IOIs at phrase boundaries with others fir each participant.

```{r ioi_dyn_ch, echo = FALSE}
# Define phrase boundaries
dt_ioi_dyn$Boundary <- "No"
dt_ioi_dyn[Interval == 8 | Interval == 16 | Interval == 24 | Interval == 41 | Interval == 49 | Interval == 57]$Boundary <- "Yes"

# For each individual
ioi_dyn_ch <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Boundary)]
ioi_dyn_ch
```

### Plot
#### Bar chdyn
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_dyn_ch_bar, echo = FALSE}
p_ioi_dyn_ch <- ggplot(data = ioi_dyn_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Boundary ~ .) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 18, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = -2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 280))
p_ioi_dyn_ch
```

## 4. Phrase boundaries - group{#dyn_group_phrase}
- Compare IOIs at phrase boundaries with others in the two conditions.
- Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations.

```{r ioi_dyn_ch_all, echo = FALSE}
# Group mean
ioi_dyn_ch_all <- ioi_dyn_ch[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_dyn_ch)/2)), by = .(Condition, Skill, Boundary)]
ioi_dyn_ch_all
```

### Plot
#### Bar chdyn
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_dyn_ch_all_bar, echo = FALSE}
p_ioi_dyn_ch_all_bar <- ggplot(data = ioi_dyn_ch_all, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width=.2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 2, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  #facet_grid(. ~ Boundary) +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 210))
p_ioi_dyn_ch_all_bar
```

#### Box plot
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_dyn_ch_all_box, echo = FALSE}
p_ioi_dyn_ch_box <- ggplot(data = ioi_dyn_ch, aes(x = Boundary, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  #annotate("text", 1.5, 188, label = "Tempo (80bpm / IOI-188ms)", vjust = 5, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics")
p_ioi_dyn_ch_box
```

### Statistics
#### Normality check
- Violate the normality
```{r normality4, echo = FALSE}
ioi_dyn_ch_norm <- by(ioi_dyn_ch$Mean, list(ioi_dyn_ch$Condition, ioi_dyn_ch$Boundary), shapiro.test)
print(ioi_dyn_ch_norm)
```
#### Two-way ANOVA (ez)
```{r stats7, echo = FALSE}
ioi_dyn_ch_aov <- ezANOVA(
  data = dt_ioi_dyn
  , dv = .(IOI)
  , wid = .(SubNr)
  , within = .(Condition, Boundary)
  , type = 3
  , detailed = TRUE
)
ioi_dyn_ch_aov$ANOVA$F <- round(ioi_dyn_ch_aov$ANOVA$F, 4)
ioi_dyn_ch_aov$ANOVA$p <- round(ioi_dyn_ch_aov$ANOVA$p, 4)
print(ioi_dyn_ch_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats8, echo = FALSE}
ioi_dyn_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_dyn_ch)
print(summary(ioi_dyn_ch_aov_2))
```

## 5. Trial by trial - individual{#dyn_ind_trial}
- Average IOIs for each trial for each participant.

```{r ioi_dyn_trial, echo = FALSE}
# For each individual
ioi_dyn_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, TrialNr)]
ioi_dyn_trial
```

### Plot
#### Line graph
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_dyn_trial_line, echo = FALSE, fig.height = 4}
p_ioi_dyn_trial <- ggplot(data = ioi_dyn_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_dyn_trial
```

## 6. Trial by trial - group{#dyn_group_trial}
- Average IOIs for each trial for each condition.

```{r ioi_dyn_trial_all, echo = FALSE}
# Group mean
ioi_dyn_trial_all <- ioi_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_dyn_trial)/2)), by = .(Condition, Skill, TrialNr)]
ioi_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi_dyn_trial_all_line, echo = FALSE}
p_ioi_dyn_trial_all <- ggplot(data = ioi_dyn_trial_all, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_dyn_trial_all
```
# Sequence plot
## Articulation{#sequence_art}
```{r seq_art, fig.width = 9, fig.height = 3, echo = FALSE}
# For each individual
ioi_art_seq <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
# Group mean
ioi_art_seq_all <- ioi_art_seq[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_art_seq)/2)), by = .(Condition, Skill, Interval)]

p_ioi_art_seq <- ggplot(data = ioi_art_seq_all, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  annotate("text", 2, 188, label = "Tempo (80bpm)", vjust = -1, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
                position = position_dodge(.05)) + 
  labs(x = "Interval", y = "IOI (ms)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,66,1))
p_ioi_art_seq
```

## Dynamics{#sequence_dyn}
```{r seq_dyn, fig.width = 9, fig.height = 3, echo = FALSE}
# For each individual
ioi_dyn_seq <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]
# Group mean
ioi_dyn_seq_all <- ioi_dyn_seq[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_dyn_seq)/2)), by = .(Condition, Skill, Interval)]

p_ioi_dyn_seq <- ggplot(data = ioi_dyn_seq_all, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  annotate("text", 2, 188, label = "Tempo (80bpm)", vjust = -1, size = 5, family = "Helvetica Neue LT Std 57 Condensed") +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
                position = position_dodge(.05)) + 
  labs(x = "Interval", y = "IOI (ms)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,66,1))
p_ioi_dyn_seq
```