### 2020-11-18 script/RNotebook_2/articulation.Rmd
update analysis scripts (check later)
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Articulation Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of all analyses for articulation.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Outline - Articulation (KOT, KOR)
## [Experimental Design](#design)

## [Articulation](#kot-articulation)

### KOT
- Individual results
  + [All](#kot_art_ind_all)
  + [Trial by trial](#kot_art_ind_trial)

- Group results and stats (if available)
  + [All](#kot_art_group_all)
  + [Trial by trial](#kot_art_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#kot_art_order_group_all)

- [Sequence plot](#sequence_art)

### KOR
- Individual results
  + [All](#kor_art_ind_all)
  + [Trial by trial](#kor_art_ind_trial)

- Group results and stats (if available)
  + [All](#kor_art_group_all)
  + [Trial by trial](#kor_art_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#kor_art_order_group_all)

## [Dynamics](#kot-dynamics)

### KOT
- Individual results
  + [All](#kot_dyn_ind_all)
  + [Trial by trial](#kot_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#kot_dyn_group_all)
  + [Trial by trial](#kot_dyn_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#kot_dyn_order_group_all)

- [Sequence plot](#sequence_dyn)

### KOR
- Individual results
  + [All](#kor_dyn_ind_all)
  + [Trial by trial](#kor_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#kor_dyn_group_all)
  + [Trial by trial](#kor_dyn_group_trial)

- Order analysis (teaching/performing first?)
  + [All](#kor_dyn_order_group_all)

## Experimental Design{#design}
- Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: legato vs. staccato / forte vs. piano)
- Dependent variables: Key-Overlap Time (KOT), Key-Overlap Ratio (KOR)

## Main predictions
Participants will exaggerate legato and staccato more while teaching.

## Excluded participnats
- SubNr 1: Most of responses (more than 90%) were removed due to outliers
- SubNr 3: Experimental error
- SubNr 8: Experimental error
- SubNr 14: Experimental error

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("ggsignif")) {install.packages("ggsignif"); require("ggsignif")}
# statistics
if (!require("stats")) {install.packages("stats"); require("stats")}
if (!require("effsize")) {install.packages("effsize"); require("effsize")}
if (!require("ez")) {install.packages("ez"); require("ez")}

# ggplots
theme_set(theme_classic())
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r file, include = FALSE}
filename_ioi = "../R/preprocessor/trimmed/data_ioi_1.txt"
filename_kot = "../R/preprocessor/trimmed/data_kot.txt"
```

```{r extract, include = FALSE}
dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed txt
dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt

# SubNr as a factor
dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
dt_kot$SubNr <- as.factor(dt_kot$SubNr)

# Exclude SubNr 1 because most responses were removed because of outliers
dt_ioi <- dt_ioi[SubNr != 1]
dt_kot <- dt_kot[SubNr != 1]

# For each individual, each block, each trial
ioi <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, BlockNr, TrialNr, Condition, Skill)]

# Calculate KOR for each interval (outliers were determined based on KOT)
dt_kor <- data.table()
for (subnr in unique(dt_kot$SubNr)){
  for (block in unique(dt_kot[SubNr == subnr]$BlockNr)){
    for (trial in unique(dt_kot[SubNr == subnr & BlockNr == block]$TrialNr)){
      dt_current <- dt_kot[SubNr == subnr & BlockNr == block & TrialNr == trial]
      dt_current$KOR <- dt_current$KOT/ioi[SubNr == subnr & ioi$BlockNr == block & TrialNr == trial]$Mean
      dt_kor <- rbind(dt_kor, dt_current)
    }
  }
}

# Include only articulation
dt_kot_art <- dt_kot[Skill == "articulation"]
dt_kor_art <- dt_kor[Skill == "articulation"]

# Include only dynamics
dt_kot_dyn <- dt_kot[Skill == "dynamics"]
dt_kor_dyn <- dt_kor[Skill == "dynamics"]
```

# Key-Overlap Time (KOT) - Articulation{#kot-articulation}
## 1. All - individual{#kot_art_ind_all}
- Average KOT for each participant.

```{r kot_art, echo = FALSE}
# For each individual
kot_art <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
kot_art
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kot_art_bar,  echo = FALSE}
p_kot_art <- ggplot(data = kot_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
p_kot_art
```

## 2. All - group{#kot_art_group_all}
- Average KOT for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

```{r kot_art_all, echo = FALSE}
# Group mean
kot_art_all <- kot_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
kot_art_all
```

### Plot
#### Bar chart
- Error bars show SEM of the mean

```{r kot_art_all_bar, echo = FALSE}　
p_kot_art_all <- ggplot(data = kot_art_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
p_kot_art_all
```

#### Box plot
```{r kot_art_all_box, echo = FALSE}
p_kot_art_all_box <- ggplot(data = kot_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
p_kot_art_all_box
```

### Statistics
#### Normality check
```{r normality1, echo = FALSE}
# Reduce unused factors
kot_art$Subcomponent <- factor(kot_art$Subcomponent)
kot_art_norm <- by(kot_art$Mean, list(kot_art$Condition, kot_art$Subcomponent), shapiro.test)
kot_art_norm
```

#### Two-way ANOVA (ez)
```{r stats1, echo = FALSE}
kot_art_aov <- ezANOVA(
  data = dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato"]
  , dv = .(KOT)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kot_art_aov$ANOVA$F <- round(kot_art_aov$ANOVA$F, 4)
kot_art_aov$ANOVA$p <- round(kot_art_aov$ANOVA$p, 4)
kot_art_aov$ANOVA
```

#### Two-way ANOVA (aov)
```{r stats2, echo = FALSE}
kot_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_art)
summary(kot_art_aov_2)
```

#### Post-hoc (direct comparison)
```{r post-hoc1, echo = FALSE}
kot_ttest_legato <- t.test(kot_art[Condition == "performing" & Subcomponent == "Legato"]$Mean, kot_art[Condition == "teaching" & Subcomponent == "Legato"]$Mean, paired = TRUE)
kot_ttest_staccato <- t.test(kot_art[Condition == "performing" & Subcomponent == "Staccato"]$Mean, kot_art[Condition == "teaching" & Subcomponent == "Staccato"]$Mean, paired = TRUE)
# # Effect size
# kot_ttest_legato_cohend <- cohen.d(kot_art$Mean[kot_art$Subcomponent == "Legato"], kot_art$Condition[kot_art$Subcomponent == "Legato"], paired = TRUE)
# kot_ttest_staccato_cohend <- cohen.d(kot_art$Mean[kot_art$Subcomponent == "Staccato"], kot_art$Condition[kot_art$Subcomponent == "Staccato"], paired = TRUE)

kot_ttest_legato
kot_ttest_staccato
#kot_ttest_legato_cohend
#kot_ttest_staccato_cohend
```

#### Post-hoc (pairwise t test)
```{r post-hoc2, echo = FALSE}
# Create a group label for a post-hoc test
kot_art$Group <- "NA"
kot_art[Condition == "performing" & Subcomponent == "Legato"]$Group <- "Performing-Legato" 
kot_art[Condition == "performing" & Subcomponent == "Staccato"]$Group <- "Performing-Staccato"
kot_art[Condition == "teaching" & Subcomponent == "Legato"]$Group <- "Teaching-Legato"
kot_art[Condition == "teaching" & Subcomponent == "Staccato"]$Group <- "Teaching-Staccato"
# pairwise comparison
kot_art_pairwise <- pairwise.t.test(kot_art$Mean, kot_art$Group, paired = TRUE, p.adjust.method = "BH")
kot_art_pairwise$p.value <- round(kot_art_pairwise$p.value, 4)
kot_art_pairwise
```

## 3. Trial by trial - individual{#kot_art_ind_trial}
- Average KOT for each trial for each participant.

```{r kot_art_trial, echo = FALSE}
# For each individual
kot_art_trial <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
kot_art_trial
```

### Plot
#### Line graph
```{r kot_art_trial_line_legato, echo = FALSE, fig.height = 4}
p_kot_art_trial_legato <- ggplot(data = kot_art_trial[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_legato
```

```{r kot_art_trial_line_staccato, echo = FALSE, fig.height = 4}
p_kot_art_trial_staccato <- ggplot(data = kot_art_trial[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_staccato
```

## 4. Trial by trial - group{#kot_art_group_trial}
- Average KOT for each trial for each condition.

```{r kot_art_trial_all, echo = FALSE}
# Group mean
kot_art_trial_all <- kot_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
kot_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r kot_art_trial_all_line_legato, echo = FALSE}
p_kot_art_trial_all_legato <- ggplot(data = kot_art_trial_all[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_all_legato
```

```{r kot_art_trial_all_line_staccato, echo = FALSE}
p_kot_art_trial_all_staccato <- ggplot(data = kot_art_trial_all[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_all_staccato
```

# Key-Overlap Ratio (KOR) - Articulation
## 1. All - individual{#kor_art_ind_all}
- Average KOR for each participant.

```{r kor_art, echo = FALSE}
# For each individual
kor_art <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
kor_art
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kor_art_bar,  echo = FALSE}
p_kor_art <- ggplot(data = kor_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOR", subtitle = "KOR: Articulation")
p_kor_art
```

## 2. All - group{#kor_art_group_all}
- Average KOR for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

```{r kor_art_all, echo = FALSE}
# Group mean
kor_art_all <- kor_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
kor_art_all
```

### Plot
#### Bar chart
- Error bars show SEM of the mean

```{r kor_art_all_bar, echo = FALSE}　
p_kor_art_all <- ggplot(data = kor_art_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KOR", subtitle = "KOR: Articulation")
p_kor_art_all
```

#### Box plot
```{r kor_art_all_box, echo = FALSE}
p_kor_art_all_box <- ggplot(data = kor_art, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOR", subtitle = "KOR: Articulation")
p_kor_art_all_box
```

### Statistics
#### Normality check
- Violate the normality
```{r normality2, echo = FALSE}
# Reduce unused factors
kor_art$Subcomponent <- factor(kor_art$Subcomponent)
kor_art_norm <- by(kor_art$Mean, list(kor_art$Condition, kor_art$Subcomponent), shapiro.test)
kor_art_norm
```

#### Two-way ANOVA (ez)
```{r stats3, echo = FALSE}
kor_art_aov <- ezANOVA(
  data = dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato"]
  , dv = .(KOR)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kor_art_aov$ANOVA$F <- round(kor_art_aov$ANOVA$F, 4)
kor_art_aov$ANOVA$p <- round(kor_art_aov$ANOVA$p, 4)
print(kor_art_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats4, echo = FALSE}
kor_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_art)
summary(kor_art_aov_2)    
```

#### Post-hoc (direct comparison)
```{r post-hoc3, echo = FALSE}
kor_ttest_legato <- t.test(kor_art[Condition == "performing" & Subcomponent == "Legato"]$Mean, kor_art[Condition == "teaching" & Subcomponent == "Legato"]$Mean, paired = TRUE)
kor_ttest_staccato <- t.test(kor_art[Condition == "performing" & Subcomponent == "Staccato"]$Mean, kor_art[Condition == "teaching" & Subcomponent == "Staccato"]$Mean, paired = TRUE)
# # Effect size
# kor_ttest_legato_cohend <- cohen.d(kor_art[Subcomponent == "Legato"]$Mean, kor_art[Subcomponent == "Legato"]$Condition, paired = TRUE)
# kor_ttest_staccato_cohend <- cohen.d(kor_art[Subcomponent == "Staccato"]$Mean, kor_art[Subcomponent == "Staccato"]$Condition, paired = TRUE)

kor_ttest_legato
kor_ttest_staccato
# kor_ttest_legato_cohend
# kor_ttest_staccato_cohend
```

#### Post-hoc (pairwise t test)
```{r post-hoc4, echo = FALSE}
# Create a group label for a post-hoc test
kor_art$Group <- "NA"
kor_art[Condition == "performing" & Subcomponent == "Legato"]$Group <- "Performing-Legato" 
kor_art[Condition == "performing" & Subcomponent == "Staccato"]$Group <- "Performing-Staccato"
kor_art[Condition == "teaching" & Subcomponent == "Legato"]$Group <- "Teaching-Legato"
kor_art[Condition == "teaching" & Subcomponent == "Staccato"]$Group <- "Teaching-Staccato"
# pairwise comparison
kor_art_pairwise <- pairwise.t.test(kor_art$Mean, kor_art$Group, paired = TRUE, p.adjust.method = "BH")
kor_art_pairwise$p.value <- round(kor_art_pairwise$p.value, 4)
kor_art_pairwise
```

## 3. Trial by trial - individual{#kor_art_ind_trial}
- Average KOR for each trial for each participant.

```{r kor_art_trial, echo = FALSE}
# For each individual
kor_art_trial <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
kor_art_trial
```

### Plot
#### Line graph
```{r kor_art_trial_line_legato, echo = FALSE, fig.height = 4}
p_kor_art_trial_legato <- ggplot(data = kor_art_trial[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_legato
```

```{r kor_art_trial_line_staccato, echo = FALSE, fig.height = 4}
p_kor_art_trial_staccato <- ggplot(data = kor_art_trial[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_staccato
```

## 4. Trial by trial - group{#kor_art_group_trial}
- Average KOR d for each trial for each condition.

```{r kor_art_trial_all, echo = FALSE}
# Group mean
kor_art_trial_all <- kor_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM =  sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
kor_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r kor_art_trial_all_line_legato, echo = FALSE}
p_kor_art_trial_all_legato <- ggplot(data = kor_art_trial_all[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_all_legato
```

```{r kor_art_trial_all_line_staccato, echo = FALSE}
p_kor_art_trial_all_staccato <- ggplot(data = kor_art_trial_all[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_all_staccato
```

# Key-Overlap Time (KOT) - Dynamics{#kot-dynamics}
## 1. All - individual{#kot_dyn_ind_all}
- Average KOT for each participant.

```{r kot_dyn, echo = FALSE}
# For each individual
kot_dyn <- dt_kot_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
kot_dyn
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kot_dyn_bar,  echo = FALSE}
p_kot_dyn <- ggplot(data = kot_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
p_kot_dyn
```

## 2. All - group{#kot_dyn_group_all}
- Average KOT for each condition.

```{r kot_dyn_all, echo = FALSE}
# Group mean
kot_dyn_all <- kot_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
kot_dyn_all
```

### Plot
#### Bar chart
- Error bars show SEM of the mean

```{r kot_dyn_all_bar, echo = FALSE}　
p_kot_dyn_all <- ggplot(data = kot_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
p_kot_dyn_all
```

#### Box plot

```{r kot_dyn_all_box, echo = FALSE}
p_kot_dyn_all_box <- ggplot(data = kot_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
p_kot_dyn_all_box
```

### Statistics
#### Normality check
- Violate the normality
```{r normality3, echo = FALSE}
# Reduce unused factors
kot_dyn$Subcomponent <- factor(kot_dyn$Subcomponent)
kot_dyn_norm <- by(kot_dyn$Mean, list(kot_dyn$Condition, kot_dyn$Subcomponent), shapiro.test)
print(kot_dyn_norm)
```

#### Two-way ANOVA (ez)
```{r stats5, echo = FALSE}
kot_dyn_aov <- ezANOVA(
  data = dt_kot_dyn[Subcomponent == "Forte" | Subcomponent == "Piano"]
  , dv = .(KOT)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kot_dyn_aov$ANOVA$F <- round(kot_dyn_aov$ANOVA$F, 4)
kot_dyn_aov$ANOVA$p <- round(kot_dyn_aov$ANOVA$p, 4)
print(kot_dyn_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats6, echo = FALSE}
kot_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_dyn)
summary(kot_dyn_aov_2)
```

## 3. Trial by trial - individual{#kot_dyn_ind_trial}
- Average KOT for each trial for each participant.

```{r kot_dyn_trial, echo = FALSE}
# For each individual
kot_dyn_trial <- dt_kot_dyn[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
kot_dyn_trial
```

### Plot
#### Line graph
```{r kot_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
p_kot_dyn_trial_forte <- ggplot(data = kot_dyn_trial[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_forte
```

```{r kot_dyn_trial_line_piano, echo = FALSE, fig.height = 4}
p_kot_dyn_trial_piano <- ggplot(data = kot_dyn_trial[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_piano
```

## 4. Trial by trial - group{#kot_dyn_group_trial}
- Average KOT for each trial for each condition.

```{r kot_dyn_trial_all, echo = FALSE}
# Group mean
kot_dyn_trial_all <- kot_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM =  sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
kot_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r kot_dyn_trial_all_line_forte, echo = FALSE}
p_kot_dyn_trial_all_forte <- ggplot(data = kot_dyn_trial_all[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_all_forte
```

```{r kot_dyn_trial_all_line_piano, echo = FALSE}
p_kot_dyn_trial_all_piano <- ggplot(data = kot_dyn_trial_all[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_all_piano
```

# Key-Overlap Ratio (KOR) - Dynamics
## 1. All - individual{#kor_dyn_ind_all}
- Average KOR for each participant.

```{r kor_dyn, echo = FALSE}
# For each individual
kor_dyn <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
kor_dyn
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kor_dyn_bar,  echo = FALSE}
p_kor_dyn <- ggplot(data = kor_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOR", subtitle = "KOR: Dynamics")
p_kor_dyn
```

## 2. All - group{#kor_dyn_group_all}
- Average KOR for each condition.

```{r kor_dyn_all, echo = FALSE}
# Group mean
kor_dyn_all <- kor_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
kor_dyn_all
```

### Plot
#### Bar chart
- Error bars show SEM of the mean

```{r kor_dyn_all_bar, echo = FALSE}　
p_kor_dyn_all <- ggplot(data = kor_dyn_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM),
                width = .2, position = position_dodge(.9)) +
  labs(y = "KOR", subtitle = "KOR: Dynamics")
p_kor_dyn_all
```

#### Box plot
```{r kor_dyn_all_box, echo = FALSE}
p_kor_dyn_all_box <- ggplot(data = kor_dyn, aes(x = Subcomponent, y = Mean, fill = Condition)) +
  geom_boxplot() +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 1, position = position_dodge(0.75)) +
  labs(y = "KOR", subtitle = "KOR: Dynamics")
p_kor_dyn_all_box
```

### Statistics
#### Normality check
- Violate the normality
```{r normality4, echo = FALSE}
# Reduce unused factors
kor_dyn$Subcomponent <- factor(kor_dyn$Subcomponent)
kor_dyn_norm <- by(kor_dyn$Mean, list(kor_dyn$Condition, kor_dyn$Subcomponent), shapiro.test)
print(kor_dyn_norm)
```

#### Two-way ANOVA (ez)
```{r stats7, echo = FALSE}
kor_dyn_aov <- ezANOVA(
  data = dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano"]
  , dv = .(KOR)
  , wid = .(SubNr)
  , within = .(Condition, Subcomponent)
  , type = 3
  , detailed = TRUE
)
kor_dyn_aov$ANOVA$F <- round(kor_dyn_aov$ANOVA$F, 4)
kor_dyn_aov$ANOVA$p <- round(kor_dyn_aov$ANOVA$p, 4)
print(kor_dyn_aov$ANOVA)
```

#### Two-way ANOVA (aov)
```{r stats8, echo = FALSE}
kor_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn)
summary(kor_dyn_aov_2)
```

## 3. Trial by trial - individual{#kor_dyn_ind_trial}
- Average KOR for each trial for each participant.

```{r kor_dyn_trial, echo = FALSE}
# For each individual
kor_dyn_trial <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
kor_dyn_trial
```

### Plot
#### Line graph
```{r kor_dyn_trial_line_forte, echo = FALSE, fig.height = 4}
p_kor_dyn_trial_forte <- ggplot(data = kor_dyn_trial[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_forte
```

```{r kor_dyn_trial_line_piano, echo = FALSE, fig.height = 4}
p_kor_dyn_trial_piano <- ggplot(data = kor_dyn_trial[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_piano
```

## 4. Trial by trial - group{#kor_dyn_group_trial}
- Average KOR for each trial for each condition.

```{r kor_dyn_trial_all, echo = FALSE}
# Group mean
kor_dyn_trial_all <- kor_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent,TrialNr)]
kor_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r kor_dyn_trial_all_line_forte, echo = FALSE}
p_kor_dyn_trial_all_forte <- ggplot(data = kor_dyn_trial_all[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_all_forte
```

```{r kor_dyn_trial_all_line_piano, echo = FALSE}
p_kor_dyn_trial_all_piano <- ggplot(data = kor_dyn_trial_all[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line(position = position_dodge(.2)) +
  geom_point(position = position_dodge(.2)) +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width = .2, position = position_dodge(.2)) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_all_piano
```

# Sequence plot
## Articulation{#sequence_art}
```{r seq_art, fig.width = 9, fig.height = 3, echo = FALSE}
# For each individual
kot_art_seq <- dt_kot_art[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Interval)]

# Group mean
kot_art_seq_all <- kot_art_seq[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Interval)]

p_kot_art_seq <- ggplot(data = kot_art_seq_all, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
                position = position_dodge(.05)) + 
  labs(x = "Interval", y = "KOT (ms)") + scale_x_continuous(breaks=seq(1,66,1))
p_kot_art_seq
```

## Dynamics{#sequence_dyn}
```{r seq_dyn, fig.width = 10, fig.height = 2, echo = FALSE}
# For each individual
kot_dyn_seq <- dt_kot_dyn[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Interval)]

# Group mean
kot_dyn_seq_all <- kot_dyn_seq[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM =  sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Interval)]

p_kot_dyn_seq <- ggplot(data = kot_dyn_seq_all, aes(x = Interval, y = Mean, group = Condition, shape = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = Mean - SEM, ymax = Mean + SEM), width=.2,
                position = position_dodge(.05)) + 
  labs(x = "Interval", y = "KOT (ms)") + scale_x_continuous(breaks=seq(1,66,1))
p_kot_dyn_seq
```