### 2020-11-25 script/R/analysis/articulation.Rmd
new analysis for kot
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Articulation Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of all analyses for articulation.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Experimental Design{#design}
- Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: legato vs. staccato / forte vs. piano)
- Dependent variables: Key-Overlap Time (KOT), Key-Overlap Ratio (KOR)

1. Main predictions
Participants will exaggerate legato and staccato more while teaching.

2. Excluded participants

- SubNr 1: Most of responses (more than 90%) were removed due to outliers
- SubNr 3: Experimental error
- SubNr 8: Experimental error
- SubNr 14: Experimental error
- SubNr 16: A deviated tempo

3. Stats packages
- [afex](https://github.com/singmann/afex) for ANOVAs
- [emmeans](https://github.com/rvlenth/emmeans) for posthoc comparisons

# Outline - Articulation (KOT, KOR)
## [Articulation](#kot-articulation)

### KOT
- Individual results
  + [All](#kot_art_ind_all)
  + [Trial by trial](#kot_art_ind_trial)

- Group results and stats (if available)
  + [All](#kot_art_group_all)
  + [Trial by trial](#kot_art_group_trial)

- [Sequence plot](#sequence_art)

### KOR
- Individual results
  + [All](#kor_art_ind_all)
  + [Trial by trial](#kor_art_ind_trial)

- Group results and stats (if available)
  + [All](#kor_art_group_all)
  + [Trial by trial](#kor_art_group_trial)

## [Dynamics](#kot-dynamics)

### KOT
- Individual results
  + [All](#kot_dyn_ind_all)
  + [Trial by trial](#kot_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#kot_dyn_group_all)
  + [Trial by trial](#kot_dyn_group_trial)

- [Sequence plot](#sequence_dyn)

### KOR
- Individual results
  + [All](#kor_dyn_ind_all)
  + [Trial by trial](#kor_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#kor_dyn_group_all)
  + [Trial by trial](#kor_dyn_group_trial)

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
# statistics
if (!require("afex")) {install.packages("afex"); require("afex")}
if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}

# ggplots
theme_set(theme_classic()) 
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r file, include = FALSE}
filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
```

```{r extract, include = FALSE}
dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed txt
dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt

# SubNr as a factor
dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
dt_kot$SubNr <- as.factor(dt_kot$SubNr)

# For each individual, each block, each trial
ioi <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, BlockNr, TrialNr, Condition, Skill)]

# Calculate KOR for each interval (outliers were already determined based on KOT)
dt_kor <- data.table()
for (subnr in unique(dt_kot$SubNr)){
  for (block in unique(dt_kot[SubNr == subnr]$BlockNr)){
    for (trial in unique(dt_kot[SubNr == subnr & BlockNr == block]$TrialNr)){
      dt_current <- dt_kot[SubNr == subnr & BlockNr == block & TrialNr == trial]
      dt_current$KOR <- dt_current$KOT/ioi[SubNr == subnr & ioi$BlockNr == block & TrialNr == trial]$Mean
      dt_kor <- rbind(dt_kor, dt_current)
    }
  }
}

# Include only articulation
dt_kot_art <- dt_kot[Skill == "articulation"]
dt_kor_art <- dt_kor[Skill == "articulation"]

# Include only dynamics
dt_kot_dyn <- dt_kot[Skill == "dynamics"]
dt_kor_dyn <- dt_kor[Skill == "dynamics"]
```

# Key-Overlap Time (KOT) - Articulation{#kot-articulation}
## 1. All - individual{#kot_art_ind_all}
- Average KOT for each participant.

```{r kot-art, echo = FALSE}
# For each individual
kot_art <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
kot_art
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kot-art-bar,  echo = FALSE}
p_kot_art <- ggplot(data = kot_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOT (ms)", subtitle = "KOT: Articulation")
p_kot_art
```

## 2. All - group{#kot_art_group_all}
- Average KOT for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

```{r kot-art-all, echo = FALSE}
# Group mean
kot_art_all <- kot_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
kot_art_all
```

### Plot
#### Box plot
```{r kot-art-all-box, echo = FALSE}
kot_art_plot <- data.table(subject = kot_art[Condition == "teaching"]$SubNr, teaching = kot_art[Condition == "teaching"]$Mean, performing = kot_art[Condition == "performing"]$Mean, subcomponent = kot_art[Condition == "teaching"]$Subcomponent)

ggpaired(kot_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (car_aov)
```{r anova-art-kot-1, echo = FALSE}
kot_art_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_art)
summary(kot_art_aov)
```

#### Two-way ANOVA (aov)
```{r anova-art-kot-2, echo = FALSE}
kot_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_art)
summary(kot_art_aov_2)
```

#### Post-hoc (emmeans)
```{r post-hoc-1, echo = FALSE}
test(pairs(emmeans(kot_art_aov,~Condition|Subcomponent)), joint = TRUE)
pairs(emmeans(kot_art_aov,~Condition|Subcomponent), adjust = "tukey")
```

## 3. Trial by trial - individual{#kot_art_ind_trial}
- Average KOT for each trial for each participant.

```{r kot-art-trial, echo = FALSE}
# For each individual
kot_art_trial <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
kot_art_trial
```

### Plot
#### Line graph
```{r kot-art-trial-line-legato, echo = FALSE, fig.height = 4}
p_kot_art_trial_legato <- ggplot(data = kot_art_trial[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_legato
```

```{r kot-art-trial-line-staccato, echo = FALSE, fig.height = 4}
p_kot_art_trial_staccato <- ggplot(data = kot_art_trial[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_art_trial_staccato
```

## 4. Trial by trial - group{#kot_art_group_trial}
- Average KOT for each trial for each condition.

```{r kot-art-trial-all, echo = FALSE}
# Group mean
kot_art_trial_all <- kot_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
kot_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r kot-art-trial-all-line-legato, echo = FALSE}
ggline(kot_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Legato") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

```{r kot-art-trial-all-line-staccato, echo = FALSE}
ggline(kot_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Articulation/Staccato") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

# Key-Overlap Ratio (KOR) - Articulation
## 1. All - individual{#kor_art_ind_all}
- Average KOR for each participant.

```{r kor-art, echo = FALSE}
# For each individual
kor_art <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
kor_art
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kor-art-bar,  echo = FALSE}
p_kor_art <- ggplot(data = kor_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOR", subtitle = "KOR: Articulation")
p_kor_art
```

## 2. All - group{#kor_art_group_all}
- Average KOR for each condition.
- Hypothesis: Participants will exaggerate legato and staccato when teaching.

```{r kor-art-all, echo = FALSE}
# Group mean
kor_art_all <- kor_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
kor_art_all
```

### Plot
#### Bar chart
- Error bars show SEM of the mean

```{r kor-art-all-bar, echo = FALSE}
kor_art_plot <- data.table(subject = kor_art[Condition == "teaching"]$SubNr, teaching = kor_art[Condition == "teaching"]$Mean, performing = kor_art[Condition == "performing"]$Mean, subcomponent = kor_art[Condition == "teaching"]$Subcomponent)

ggpaired(kor_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (car_aov)
```{r anova-art-kor-1, echo = FALSE}
kor_art_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_art)
summary(kor_art_aov)
```

#### Two-way ANOVA (aov)
```{r anova-art-kor-2, echo = FALSE}
kor_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_art)
summary(kor_art_aov_2)    
```

#### Post-hoc (emmeans)
```{r post-hoc-2, echo = FALSE}
test(pairs(emmeans(kor_art_aov,~Condition|Subcomponent)), joint = TRUE)
pairs(emmeans(kor_art_aov,~Condition|Subcomponent), adjust = "tukey")
```

## 3. Trial by trial - individual{#kor_art_ind_trial}
- Average KOR for each trial for each participant.

```{r kor-art-trial, echo = FALSE}
# For each individual
kor_art_trial <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
kor_art_trial
```

### Plot
#### Line graph
```{r kor-art-trial-line-legato, echo = FALSE, fig.height = 4}
p_kor_art_trial_legato <- ggplot(data = kor_art_trial[Subcomponent == "Legato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Legato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_legato
```

```{r kor-art-trial-line-staccato, echo = FALSE, fig.height = 4}
p_kor_art_trial_staccato <- ggplot(data = kor_art_trial[Subcomponent == "Staccato"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Articulation/Staccato") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_art_trial_staccato
```

## 4. Trial by trial - group{#kor_art_group_trial}
- Average KOR d for each trial for each condition.

```{r kor-art-trial-all, echo = FALSE}
# Group mean
kor_art_trial_all <- kor_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM =  sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
kor_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r kor-art-trial-all-line-legato, echo = FALSE}
ggline(kor_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

```{r kor-art-trial-all-line-staccato, echo = FALSE}
ggline(kor_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Articulation/Legato") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

# Key-Overlap Time (KOT) - Dynamics{#kot-dynamics}
## 1. All - individual{#kot_dyn_ind_all}
- Average KOT for each participant.

```{r kot-dyn, echo = FALSE}
# For each individual
kot_dyn <- dt_kot_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
kot_dyn
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kot-dyn-bar,  echo = FALSE}
p_kot_dyn <- ggplot(data = kot_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOT (ms)", subtitle = "KOT: Dynamics")
p_kot_dyn
```

## 2. All - group{#kot_dyn_group_all}
- Average KOT for each condition.

```{r kot-dyn-all, echo = FALSE}
# Group mean
kot_dyn_all <- kot_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
kot_dyn_all
```

### Plot
#### Box plot

```{r kot-dyn-all-box, echo = FALSE}
kot_dyn_plot <- data.table(subject = kot_dyn[Condition == "teaching"]$SubNr, teaching = kot_dyn[Condition == "teaching"]$Mean, performing = kot_dyn[Condition == "performing"]$Mean, subcomponent = kot_dyn[Condition == "teaching"]$Subcomponent)

ggpaired(kot_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Dynamics") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (car_aov)
```{r anova-dyn-kot-1, echo = FALSE}
kot_dyn_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_dyn)
summary(kot_dyn_aov)
```

#### Two-way ANOVA (aov)
```{r anova-dyn-kot-2, echo = FALSE}
kot_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kot_dyn)
summary(kot_dyn_aov_2)
```

## 3. Trial by trial - individual{#kot_dyn_ind_trial}
- Average KOT for each trial for each participant.

```{r kot-dyn-trial, echo = FALSE}
# For each individual
kot_dyn_trial <- dt_kot_dyn[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
kot_dyn_trial
```

### Plot
#### Line graph
```{r kot-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
p_kot_dyn_trial_forte <- ggplot(data = kot_dyn_trial[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_forte
```

```{r kot-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
p_kot_dyn_trial_piano <- ggplot(data = kot_dyn_trial[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOT (ms)", subtitle = "KOT: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_kot_dyn_trial_piano
```

## 4. Trial by trial - group{#kot_dyn_group_trial}
- Average KOT for each trial for each condition.

```{r kot-dyn-trial-all, echo = FALSE}
# Group mean
kot_dyn_trial_all <- kot_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM =  sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
kot_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r kot-dyn-trial-all-line-forte, echo = FALSE}
ggline(kot_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Forte") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

```{r kot-dyn-trial-all-line-piano, echo = FALSE}
ggline(kot_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOT (ms)", title = "KOT: Dynamics/Piano") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

# Key-Overlap Ratio (KOR) - Dynamics
## 1. All - individual{#kor_dyn_ind_all}
- Average KOR for each participant.

```{r kor-dyn, echo = FALSE}
# For each individual
kor_dyn <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
kor_dyn
```

### Plot
#### Bar chart
- Error bars show SD of the mean

```{r kor-dyn-bar,  echo = FALSE}
p_kor_dyn <- ggplot(data = kor_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Subcomponent ~ .) +
  labs(y = "KOR", subtitle = "KOR: Dynamics")
p_kor_dyn
```

## 2. All - group{#kor_dyn_group_all}
- Average KOR for each condition.

```{r kor-dyn-all, echo = FALSE}
# Group mean
kor_dyn_all <- kor_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
kor_dyn_all
```

### Plot
#### Box plot
```{r kor-dyn-all-box, echo = FALSE}
kor_dyn_plot <- data.table(subject = kor_dyn[Condition == "teaching"]$SubNr, teaching = kor_dyn[Condition == "teaching"]$Mean, performing = kor_dyn[Condition == "performing"]$Mean, subcomponent = kor_dyn[Condition == "teaching"]$Subcomponent)

ggpaired(kor_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (car_aov)
```{r anova-dyn-kor-1, echo = FALSE}
kor_dyn_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn)
summary(kor_dyn_aov)
```

#### Two-way ANOVA (aov)
```{r anova-dyn-kor-2, echo = FALSE}
kor_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = kor_dyn)
summary(kor_dyn_aov_2)
```

## 3. Trial by trial - individual{#kor_dyn_ind_trial}
- Average KOR for each trial for each participant.

```{r kor-dyn-trial, echo = FALSE}
# For each individual
kor_dyn_trial <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
kor_dyn_trial
```

### Plot
#### Line graph
```{r kor-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
p_kor_dyn_trial_forte <- ggplot(data = kor_dyn_trial[Subcomponent == "Forte"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Forte") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_forte
```

```{r kor-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
p_kor_dyn_trial_piano <- ggplot(data = kor_dyn_trial[Subcomponent == "Piano"], aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "KOR", subtitle = "KOR: Dynamics/Piano") + scale_x_continuous(breaks=seq(1,8,1))
p_kor_dyn_trial_piano
```

## 4. Trial by trial - group{#kor_dyn_group_trial}
- Average KOR for each trial for each condition.

```{r kor-dyn-trial-all, echo = FALSE}
# Group mean
kor_dyn_trial_all <- kor_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent,TrialNr)]
kor_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean

```{r kor-dyn-trial-all-line-forte, echo = FALSE}
ggline(kor_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Forte") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

```{r kor_dyn_trial_all_line_piano, echo = FALSE}
ggline(kor_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "KOR", title = "KOR: Dynamics/Piano") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

# Sequence plot
## Articulation{#sequence_art}
```{r seq-art, fig.width = 10, fig.height = 2, echo = FALSE}
# For each individual
kot_art_seq <- dt_kot_art[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Interval)]

ggline(kot_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Articulation") +
  scale_x_continuous(breaks=seq(1,66,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

## Dynamics{#sequence_dyn}
```{r seq-dyn, fig.width = 10, fig.height = 2, echo = FALSE}
# For each individual
kot_dyn_seq <- dt_kot_dyn[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Interval)]

ggline(kot_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "KOT (ms)", title = "KOT: Dynamics") +
  scale_x_continuous(breaks=seq(1,66,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```