### 2020-11-25 script/R/analysis/ioi.Rmd
correct some comments
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Tempo Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of tempo analysis.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Experimental Design{#design}
- Independent variables: Condition (2 levels: teaching vs. performing) x Technique (2 levels: articulation vs. dynamics)
- Dependent variables: Interonset intervals (IOIs)

1. Main predictions
- Participants will play slower while teaching regardless of the techniques.

2. Excluded participants

- SubNr 1: Most of responses (more than 90%) were removed due to outliers
- SubNr 3: Experimental error
- SubNr 8: Experimental error
- SubNr 14: Experimental error
- SubNr 16: A deviated tempo

3. Stats packages
- [afex](https://github.com/singmann/afex) for ANOVAs

# Outline - Tempo (IOIs)
## [Articulation](#ioi-articulation)

- Individual results
  + [All](#art_ind_all)
  + [Phrase bondaries](#art_ind_phrase)
  + [Trial by trial](#art_ind_trial)

- Group results and stats (if available)
  + [All](#art_group_all)
  + [Phrase bondaries](#art_group_phrase)
  + [Trial by trial](#art_group_trial)

- [Sequence plot](#sequence_art)

## [Dynamics](#ioi-dynamics)

- Individual results
  + [All](#dyn_ind_all)
  + [Phrase bondaries](#dyn_ind_phrase)
  + [Trial by trial](#dyn_ind_trial)

- Group results and stats (if available)
  + [All](#dyn_group_all)
  + [Phrase bondaries](#dyn_group_phrase)
  + [Trial by trial](#dyn_group_trial)

- [Sequence plot](#sequence_dyn)

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
# statistics
if (!require("afex")) {install.packages("afex"); require("afex")}

# ggplots
theme_set(theme_classic()) 
theme_update(text = element_text(size = 20, family = "Helvetica Neue LT Std 57 Condensed"), legend.position = "bottom")
```

```{r file, include = FALSE}
filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
```

```{r extract, include = FALSE}
dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv

# SubNr as a factor
dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)

# Include only articulation
dt_ioi_art <- dt_ioi[Skill == "articulation"]
# Include only dynamics
dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
```

# IOI-Articulation 
## 1. All - individual{#art_ind_all}
- Average IOIs for each participant.

```{r ioi-art, echo = FALSE}
# For each individual
ioi_art <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
ioi_art
```

### Plot
#### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-art-bar,  echo = FALSE}
p_ioi_art <- ggplot(data = ioi_art, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 280))
p_ioi_art
```

## 2. All - group{#art_group_all}
- Average IOIs for each condition.
- Hypothesis: Participants will play slower when teaching.

```{r ioi-art-all, echo = FALSE}
# Group mean
ioi_art_all <- ioi_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill)]
ioi_art_all
```

### Plot
#### Box plot
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-art-plot, echo = FALSE}
ioi_art_plot <- data.table(subject = ioi_art[Condition == "teaching"]$SubNr, teaching = ioi_art[Condition == "teaching"]$Mean, performing = ioi_art[Condition == "performing"]$Mean)

ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
  geom_hline(yintercept = 188, linetype = "dashed") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "none")
```

### Statistics
#### Normality check
- Violate the normality for differences between the two conditions

```{r normality1, echo = FALSE}
# Normality check
ioi_art_diff <- ioi_art[Condition == "teaching"]$Mean-ioi_art[Condition == "performing"]$Mean
ioi_art_norm <- shapiro.test(ioi_art_diff)
ioi_art_norm
```

#### Wilcoxon signed-rank test
```{r wilcoxon-art, echo = FALSE}
ioi_art_wilcoxon <- wilcox.test(Mean ~ Condition, data = ioi_art, paired = TRUE, alternative = "two.sided")
ioi_art_wilcoxon
```

## 3. Phrase boundaries - individual{#art_ind_phrase}
- Compare IOIs at phrase boundaries with others for each participant.

```{r ioi-art-ch, echo = FALSE}
# Define phrase boundaries
dt_ioi_art$Boundary <- "No"
dt_ioi_art[Subcomponent == "LtoS" | Subcomponent == "StoL"]$Boundary <- "Yes"

# For each individual
ioi_art_ch <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Boundary)]
ioi_art_ch
```

### Plot
#### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-art-ch-bar, echo = FALSE}
p_ioi_art_ch <- ggplot(data = ioi_art_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Boundary ~ .) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  labs(y = "IOI (ms)", subtitle = "IOI: Articulation") + coord_cartesian(ylim = c(100, 280))
p_ioi_art_ch
```

## 4. Phrase boundaries - group{#art_group_phrase}
- Compare IOIs at phrase boundaries with others in the two conditions.
- Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations (especially for teaching purposes).

```{r ioi-art-ch-all, echo = FALSE}
# Group mean
ioi_art_ch_all <- ioi_art_ch[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(nrow(ioi_art_ch)/2)), by = .(Condition, Skill, Boundary)]
ioi_art_ch_all
```

### Plot
#### Box plot
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-art-ch-all-bar, echo = FALSE}
ioi_art_ch_plot <- data.table(subject = ioi_art_ch[Condition == "teaching"]$SubNr, teaching = ioi_art_ch[Condition == "teaching"]$Mean, performing = ioi_art_ch[Condition == "performing"]$Mean, boundary = ioi_art_ch[Condition == "teaching"]$Boundary)

ggpaired(ioi_art_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") +
  geom_hline(yintercept = 188, linetype = "dashed") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (aov_car)
```{r anova-art-boundary, echo = FALSE}
ioi_art_ch_aov <- aov_car(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_art_ch)
summary(ioi_art_ch_aov)
```

#### Two-way ANOVA (aov)
```{r anova-art-boundary-2, echo = FALSE}
ioi_art_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_art_ch)
summary(ioi_art_ch_aov_2)
```

## 5. Trial by trial - individual{#art_ind_trial}
- Average IOIs for each trial for each participant.

```{r ioi-art-trial, echo = FALSE}
# For each individual
ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, TrialNr)]
ioi_art_trial
```

### Plot
#### Line graph
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-art-trial-line, echo = FALSE, fig.height = 4}
p_ioi_art_trial <- ggplot(data = ioi_art_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Articulation") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_art_trial
```

## 6. Trial by trial - group{#art_group_trial}
- Average IOIs for each trial for each condition.

```{r ioi-art-trial-all, echo = FALSE}
# Group mean
ioi_art_trial_all <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, TrialNr)]
ioi_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-art-trial-all-line, echo = FALSE}
ggline(ioi_art_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Articulation") +
  geom_hline(yintercept = 188, linetype = "dashed") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

# IOI-Dynamics
## 1. All - individual{#dyn_ind_all}
- Average IOIs for each participant.

```{r ioi-dyn, echo = FALSE}
# For each individual
ioi_dyn <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
ioi_dyn
```

### Plot
#### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-dyn-bar,  echo = FALSE}
p_ioi_dyn <- ggplot(data = ioi_dyn, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 280))
p_ioi_dyn
```

## 2. All - group{#dyn_group_all}
- Average IOIs for each condition.
- Hypothesis: Participants will play slower when teaching.

```{r ioi-dyn-all, echo = FALSE}
# Group mean
ioi_dyn_all <- ioi_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean),  SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill)]
ioi_dyn_all
```

### Plot
#### Box plot
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-dyn-all-box, echo = FALSE}
ioi_dyn_plot <- data.table(subject = ioi_dyn[Condition == "teaching"]$SubNr, teaching = ioi_dyn[Condition == "teaching"]$Mean, performing = ioi_dyn[Condition == "performing"]$Mean)

ggpaired(ioi_dyn_plot, cond1 = "teaching", cond2 = "performing", color = "condition", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
  geom_hline(yintercept = 188, linetype = "dashed") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "none")
```

### Statistics
#### Normality check
- Violate the normality for differences between the two conditions

```{r normality2, echo = FALSE}
# Normality check
ioi_dyn_diff <- ioi_dyn[Condition == "teaching"]$Mean-ioi_dyn[Condition == "performing"]$Mean
ioi_dyn_norm <- shapiro.test(ioi_dyn_diff)
ioi_dyn_norm
```

#### Wilcoxon signed-rank test
```{r wilcoxon-dyn, echo = FALSE}
ioi_dyn_wilcoxon <- wilcox.test(Mean ~ Condition, data = ioi_dyn, paired = TRUE, alternative = "two.sided")
ioi_dyn_wilcoxon
```

## 3. Phrase boundaries - individual{#dyn_ind_phrase}
- Compare IOIs at phrase boundaries with others for each participant.

```{r ioi-dyn-ch, echo = FALSE}
# Define phrase boundaries
dt_ioi_dyn$Boundary <- "No"
dt_ioi_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF"]$Boundary <- "Yes"

# For each individual
ioi_dyn_ch <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Boundary)]
ioi_dyn_ch
```

### Plot
#### Bar chart
- Error bars show SD of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-dyn-ch-bar, echo = FALSE}
p_ioi_dyn_ch <- ggplot(data = ioi_dyn_ch, aes(x = SubNr, y = Mean, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                width = .2, position = position_dodge(.9)) +
  facet_grid(Boundary ~ .) +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  labs(y = "IOI (ms)", subtitle = "IOI: Dynamics") + coord_cartesian(ylim = c(100, 280))
p_ioi_dyn_ch
```

## 4. Phrase boundaries - group{#dyn_group_phrase}
- Compare IOIs at phrase boundaries with others in the two conditions.
- Hypothesis: Participants will play more slowly at phrase boundaries compared to other locations (especially for teaching purposes).

```{r ioi-dyn-ch-all, echo = FALSE}
# Group mean
ioi_dyn_ch_all <- ioi_dyn_ch[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Boundary)]
ioi_dyn_ch_all
```

### Plot
#### Box plot
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-dyn-ch-all-bar, echo = FALSE}
ioi_dyn_ch_plot <- data.table(subject = ioi_dyn_ch[Condition == "teaching"]$SubNr, teaching = ioi_dyn_ch[Condition == "teaching"]$Mean, performing = ioi_dyn_ch[Condition == "performing"]$Mean, boundary = ioi_dyn_ch[Condition == "teaching"]$Boundary)

ggpaired(ioi_dyn_ch_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "boundary", line.size = 0.3, line.color = "gray", palette = "aaas", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
  geom_hline(yintercept = 188, linetype = "dashed") +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed", ) +
  theme(legend.position = "none")
```

### Statistics
#### Two-way ANOVA (aov_car)
```{r anova-dyn-boundary, echo = FALSE}
ioi_dyn_ch_aov <- aov_car(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_dyn_ch)
summary(ioi_dyn_ch_aov)
```

#### Two-way ANOVA (aov)
```{r anova-art-boundary2, echo = FALSE}
ioi_dyn_ch_aov_2 <- aov(Mean ~ Condition*Boundary + Error(SubNr/(Condition*Boundary)), data = ioi_dyn_ch)
summary(ioi_dyn_ch_aov_2)
```

## 5. Trial by trial - individual{#dyn_ind_trial}
- Average IOIs for each trial for each participant.

```{r ioi-dyn-trial, echo = FALSE}
# For each individual
ioi_dyn_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, TrialNr)]
ioi_dyn_trial
```

### Plot
#### Line graph
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-dyn-trial-line, echo = FALSE, fig.height = 4}
p_ioi_dyn_trial <- ggplot(data = ioi_dyn_trial, aes(x = TrialNr, y = Mean, group = Condition, colour = Condition)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept = 188, linetype = "dashed") + # Tempo
  facet_wrap(SubNr ~ .) + 
  labs(x = "Trial", y = "IOI (ms)", subtitle = "IOI: Dynamics") + scale_x_continuous(breaks=seq(1,8,1))
p_ioi_dyn_trial
```

## 6. Trial by trial - group{#dyn_group_trial}
- Average IOIs for each trial for each condition.

```{r ioi-dyn-trial-all, echo = FALSE}
# Group mean
ioi_dyn_trial_all <- ioi_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean),  SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, TrialNr)]
ioi_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms)

```{r ioi-dyn-trial-all-line, echo = FALSE}
ggline(ioi_dyn_trial, x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Trial", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
  geom_hline(yintercept = 188, linetype = "dashed") +
  scale_x_continuous(breaks=seq(1,8,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```
# Sequence plot
## Articulation{#sequence_art}
```{r seq-art, fig.width = 10, fig.height = 2, echo = FALSE}
# For each individual
ioi_art_seq <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]

ggline(ioi_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Articulation") +
  geom_hline(yintercept = 188, linetype = "dashed") +
  scale_x_continuous(breaks=seq(1,66,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```

## Dynamics{#sequence_dyn}
```{r seq-dyn, fig.width = 10, fig.height = 2, echo = FALSE}
# For each individual
ioi_dyn_seq <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, Interval)]

ggline(ioi_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", palette = "aaas", xlab = "Interval", ylab = "IOIs (ms)", title = "IOI: Dynamics") +
  geom_hline(yintercept = 188, linetype = "dashed") +
  scale_x_continuous(breaks=seq(1,66,1)) +
  theme_pubr(base_size = 20, base_family = "Helvetica Neue LT Std 57 Condensed") +
  theme(legend.position = "bottom")
```