### 2020-12-02 script/R/demographics/questionnaire.Rmd
update demographics
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Demographics: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of demographic information

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
```

# Questionnaire

```{r file, include = FALSE}
filename_q = "./questionnaire.csv"
```

```{r extract, echo = FALSE}
# read csv files
dt <- fread(filename_q, header = T, sep = ",", dec = ".", na.string = "NA")

# change some colnames
colnames(dt)[2] <- "SubNr"
colnames(dt)[3] <- "Age"
colnames(dt)[4] <- "Gender"
colnames(dt)[7] <- "Handedness"

# change some characteristics
dt$Age <- as.numeric(dt$Age)

# extract the total number of years of training in piano (input manually based on the original csv file)
dt$PianoTotalPractice <- c(14, 20, 27, 21, 10, 16, 9, 13, 25, 14, 13, 11, 6, 4, 6, 8, 21, 15, 18, 7, 11, 12, 12, 18, 12, 5, 5, 10, 16, 10, 8, 22, 10, 5, 4, 14)

# exclude participants
dt_included <- dt[SubNr != 1 & SubNr != 3 & SubNr != 8 & SubNr != 14 & SubNr != 16]

print(dt_included)
```

## Descriptive stats for included participants
- Number of included participants: `r nrow(dt_included)` (mean age: `r round(mean(dt_included$Age), 2)`, sd: `r round(sd(dt_included$Age), 2)`)
- Male: `r nrow(dt_included[Gender == "Male"])` (mean age: `r round(mean(dt_included[Gender == "Male"]$Age), 2)`, sd: `r round(sd(dt_included[Gender == "Male"]$Age), 2)`)
- Female: `r nrow(dt_included[Gender == "Female"])` (mean age: `r round(mean(dt_included[Gender == "Female"]$Age), 2)`, sd: `r round(sd(dt_included[Gender == "Female"]$Age), 2)`)
- Average training years: `r round(mean(dt_included$PianoTotalPractice), 2)` years
- Handedness: Right (`r nrow(dt_included[Handedness == "Right"])`), Left (`r nrow(dt_included[Handedness == "Left"])`), Both (`r nrow(dt_included[Handedness == "Both"])`)

## Some descriptive questions
1. Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?

```{r 1, echo = FALSE}
data.table("Answer" = dt_included$`Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?`)
```

2. Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.

```{r 2, echo = FALSE}
data.table("Answer" = dt_included$`Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.`)
```