### 2020-12-02 script/analysis/preprocessor/error_summary.Rmd
rename folders
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Error Summary (teaching-v1.0)"
output: html_notebook
---

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Workflow
1. filter with `filtering.R`
2. check the details of errors and remove some participants if needed (**this script**: `error_summary.Rmd`)
3. trim data (i.e., remove outliers) with `trimming.R`
4. check trimmed data and remove some participants if needed (`included_summary.Rmd`)

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("here")) {install.packages("here"); require("here")}

# set working directory
here::i_am("error_summary.Rmd")

# function
source(here("function.R"))

# read a text file for ideal performance
dt_ideal <- read.table(here("ideal.txt"))
colnames(dt_ideal) <- "Pitch"
dt_ideal$RowNr <- 1:nrow(dt_ideal)
setcolorder(dt_ideal, c(2, 1))
# 
dt_onset <- fread(file = here("filtered", "dt_correct_onset.txt"), sep = ",")
dt_offset <- fread(file = here("filtered", "dt_correct_offset.txt"), sep = ",")
error_onset <- fread(file = here("filtered", "error_onset.txt"), sep = ",")
error_offset <- fread(file = here("filtered", "error_offset.txt"), sep = ",")
```

# Design
- Each performance (1 trial) contains 67 responses for onsets and 67 responses for offsets (i.e., if there is no mistake, 134 responses in total)
- Each block contains 8 trials and there are 4 blocks (teaching vs. performing x articulation vs. dynamics)
- Note: SubNr 8 was excluded due to an experimental error

# Outline
1. [Onset reports](#onset)
2. [Offset reports](#offset)

## ONSET
### Corrected Error Onset (All)
```{r onset-all, echo = FALSE}
error_onset_all <- error_onset[CorrectionNr != "Exclude",]
error_onset_all
```

### Corrected Error Onset (Summary)
```{r onset-summary, echo = FALSE}
error_onset_all[startsWith(Reason, "Substituted")]$Reason <- "Substituted"
error_onset_summary <- error_onset_all[, .(N = .N, Sum = sum(as.numeric(CorrectionNr))), by = .(Reason)]
error_onset_summary
```

### Missing/Excluded Trials Onset (All)
```{r na-onset-all, echo = FALSE}
missing_onset <- checker(dt_onset, dt_ideal)
missing_onset
```

1. **2 participants should be removed from the entire data analysis because they missed at least one block (16 trials)**

- SubNr 3: experimental error (8 trials)
- SubNr 14: played in octave higher (8 trials)

2. **some performances were removed because 1) participants didn't follow sheet music or 2) participants didn't finish till the end (5 trials)**

### Missing/Excluded Trials Onset (Summary)
```{r na-onset-summary, echo = FALSE}
missing_onset_summary <- missing_onset[, .(N = .N), by = .(Reason)]
missing_onset_summary
```

## OFFSET
### Corrected Error Offset (All)
```{r offset-all, echo = FALSE}
error_offset_all <- error_offset[CorrectionNr != "Exclude",]

# separate errors depending on reasons
# SubNr 11, BlockNr 4, TrialNr 8
error_offset_all <- rbind(error_offset_all, error_offset_all[4])
error_offset_all$Reason[4] <- "Extra Notes"
error_offset_all$CorrectionNr[4] <- 1
error_offset_all$Reason[nrow(error_offset_all)] <- "Substituted"
error_offset_all$CorrectionNr[nrow(error_offset_all)] <- 2
# SubNr 27, BlockNr 3, TrialNr 6
error_offset_all <- rbind(error_offset_all, error_offset_all[5])
error_offset_all$Reason[5] <- "Extra Notes"
error_offset_all$CorrectionNr[5] <- 1
error_offset_all$Reason[nrow(error_offset_all)] <- "Substituted"
error_offset_all$CorrectionNr[nrow(error_offset_all)] <- 6

error_offset_all
```

### Corrected Error Offset (Summary)
```{r offset-summary, echo = FALSE}
error_offset_all[startsWith(Reason, "Substituted")]$Reason <- "Substituted"
error_offset_summary <- error_offset_all[, .(N = .N, Sum = sum(as.numeric(CorrectionNr))), by = .(Reason)]
error_offset_summary
```

### Missing/Excluded Trials Offset (All)
```{r na-offset-all, echo = FALSE}
missing_offset <- checker(dt_offset, dt_ideal)
missing_offset
```

1. **2 participants should be removed from the entire data analysis because they missed at least one block (16 trials)**

- SubNr 3: experimental error (8 trials)
- SubNr 14: played in octave higher (8 trials)

2. **some performances were removed because 1) participants didn't follow sheet music or 2) participants didn't finish till the end (5 trials)**

### Missing/Excluded Trials Offset (Summary)
```{r na-offset-summary, echo = FALSE}
missing_offset_summary <- missing_offset[, .(N = .N), by = .(Reason)]
missing_offset_summary
```