### 2020-12-02 script/R/preprocessor/error_summary.Rmd
removed old analysis and update error_summary
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Error Summary (teaching-v2.0)"
output: html_notebook
---

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Workflow
1. filter with `filtering.R`
2. check the details of errors and remove some participants if needed (**this script**: `error_summary.Rmd`)
3. trim data (i.e., remove outliers) with `trimming.R`
4. check trimmed data and remove some participants if needed (`included_summary.Rmd`)

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("here")) {install.packages("here"); require("here")}

# set working directory
here::i_am("error_summary.Rmd")

# function
source(here("function.R"))

# read a text file for ideal performance
dt_ideal <- read.table(here("ideal.txt"))
colnames(dt_ideal) <- "Pitch"
dt_ideal$RowNr <- 1:nrow(dt_ideal)
setcolorder(dt_ideal, c(2, 1))
# 
dt_onset <- fread(file = here("filtered", "dt_correct_onset.txt"), sep = ",")
dt_offset <- fread(file = here("filtered", "dt_correct_offset.txt"), sep = ",")
error_onset <- fread(file = here("filtered", "error_onset.txt"), sep = ",")
error_offset <- fread(file = here("filtered", "error_offset.txt"), sep = ",")
```

# Design
- Each performance (1 trial) contains 72 responses for onsets and 72 responses for offsets (i.e., if there is no mistake, 144 responses in total)
- Each block contains 8 trials and there are 4 blocks (teaching vs. performing x articulation vs. dynamics)
- Note: SubNr 3 was excluded due to insufficient motor skills

# Outline
1. [Onset reports](#onset)
2. [Offset reports](#offset)

## ONSET
### Corrected Error Onset (All)
```{r onset-all, echo = FALSE}
error_onset_all <- error_onset[CorrectionNr != "Exclude"]

# separate errors depending on reasons
# SubNr 10, BlockNr 1, TrialNr 6
error_onset_all <- rbind(error_onset_all, error_onset_all[2])
error_onset_all$Reason[2] <- "Extra Notes"
error_onset_all$CorrectionNr[2] <- 2
error_onset_all$Reason[nrow(error_onset_all)] <- "Substituted"
error_onset_all$CorrectionNr[nrow(error_onset_all)] <- 1

error_onset_all
```

### Corrected Error Onset (Summary)
```{r onset-summary, echo = FALSE}
error_onset_all[startsWith(Reason, "Substituted")]$Reason <- "Substituted"
error_onset_summary <- error_onset_all[, .(N = .N, Sum = sum(as.numeric(CorrectionNr))), by = .(Reason)]
error_onset_summary
```

### Missing/Excluded Trials Onset (All)
```{r na-onset-all, echo = FALSE}
missing_onset <- checker(dt_onset, dt_ideal)
missing_onset
```

### Missing/Excluded Trials Onset (Summary)
```{r na-onset-summary, echo = FALSE}
missing_onset_summary <- missing_onset[, .(N = .N), by = .(Reason)]
missing_onset_summary
```

## OFFSET
### Corrected Error Offset (All)
```{r offset-all, echo = FALSE}
error_offset_all <- error_offset[CorrectionNr != "Exclude",]

# separate errors depending on reasons
# SubNr 1, BlockNr 2, TrialNr 3
error_offset_all <- rbind(error_offset_all, error_offset_all[1])
error_offset_all$Reason[1] <- "Extra Notes"
error_offset_all$CorrectionNr[1] <- 1
error_offset_all$Reason[nrow(error_offset_all)] <- "Substituted"
error_offset_all$CorrectionNr[nrow(error_offset_all)] <- 4
# SubNr 8, BlockNr 4, TrialNr 2
error_offset_all <- rbind(error_offset_all, error_offset_all[5])
error_offset_all$Reason[5] <- "Extra Notes"
error_offset_all$CorrectionNr[5] <- 1
error_offset_all$Reason[nrow(error_offset_all)] <- "Substituted"
error_offset_all$CorrectionNr[nrow(error_offset_all)] <- 2
# SubNr 10, BlockNr 1, TrialNr 6
error_offset_all <- rbind(error_offset_all, error_offset_all[6])
error_offset_all$Reason[6] <- "Extra Notes"
error_offset_all$CorrectionNr[6] <- 2
error_offset_all$Reason[nrow(error_offset_all)] <- "Substituted"
error_offset_all$CorrectionNr[nrow(error_offset_all)] <- 1
# SubNr 15, BlockNr 3, TrialNr 5
error_offset_all <- rbind(error_offset_all, error_offset_all[7])
error_offset_all$Reason[7] <- "Extra Notes"
error_offset_all$CorrectionNr[7] <- 1
error_offset_all$Reason[nrow(error_offset_all)] <- "Substituted"
error_offset_all$CorrectionNr[nrow(error_offset_all)] <- 2

error_offset_all
```

### Corrected Error Offset (Summary)
```{r offset-summary, echo = FALSE}
error_offset_all[startsWith(Reason, "Substituted")]$Reason <- "Substituted"
error_offset_summary <- error_offset_all[, .(N = .N, Sum = sum(as.numeric(CorrectionNr))), by = .(Reason)]
error_offset_summary
```

### Missing/Excluded Trials Offset (All)
```{r na-offset-all, echo = FALSE}
missing_offset <- checker(dt_offset, dt_ideal)
missing_offset
```

### Missing/Excluded Trials Offset (Summary)
```{r na-offset-summary, echo = FALSE}
missing_offset_summary <- missing_offset[, .(N = .N), by = .(Reason)]
missing_offset_summary
```