### 2020-12-02 analysis/preprocessor/included_summary.Rmd
rename folders
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Included  Summary (teaching-v2.0)"
output: html_notebook
---

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Workflow
1. filter with `filtering.R`
2. check the details of errors and remove some participants if needed (`error_summary.Rmd`)
3. trim data (i.e., remove outliers) with `trimming_ioi.R`, `trimming_kot_kor.R`, `trimming_vel.R`
4. check trimmed data and remove some participants if needed (**this script**: `included_summary.Rmd`)

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("here")) {install.packages("here"); require("here")}

# set working directory
here::i_am("included_summary.Rmd")

# read txt files
dt_ioi <- fread(file = here("trimmed", "data_ioi_1.txt"), sep = ",")
dt_kor <- fread(file = here("trimmed", "data_kor.txt"), sep = ",")
dt_kot <- fread(file = here("trimmed", "data_kot.txt"), sep = ",")
dt_vel <- fread(file = here("trimmed", "data_vel.txt"), sep = ",")
dt_vel_diff <- fread(file = here("trimmed", "data_vel_diff.txt"), sep = ",")
```

# Design
- Each performance (1 trial) contains 72 responses for onsets and 72 responses for offsets (i.e., if there is no mistake, 144 responses in total)
- Each block contains 8 trials and there are 4 blocks (teaching vs. performing x articulation vs. dynamics)
- Note: SubNr 3 was excluded due to insufficient motor skills

### IOIs (All)
```{r ioi, echo = FALSE}
dt_ioi
```

### IOIs (Summary)
```{r ioi-summary, echo = FALSE}
dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill)]
```

### KOR (All)
```{r kor, echo = FALSE}
dt_kor
```

### KOR (Summary)
```{r kor-summary, echo = FALSE}
dt_kor[, .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, Subcomponent)]
```

### KOT (All)
```{r kot, echo = FALSE}
dt_kot
```

### KOT (Summary)
```{r kot-summary, echo = FALSE}
dt_kot[, .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, Subcomponent)]
```

### Key Velocity (All)
```{r vel, echo = FALSE}
dt_vel
```

### Key Velocity (Summary)
```{r vel-summary, echo = FALSE}
dt_vel[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
```

### KV Difference (All)
```{r vel-diff, echo = FALSE}
dt_vel_diff
```

### KV Difference (Summary)
```{r vel-diff-summary, echo = FALSE}
dt_vel_diff[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
```

## Excluded participnats
- SubNr 3: Insufficient motor skills

### Export only included participants
data_analysis~.txt are used for data analysis.
```{r export, echo = TRUE}
fwrite(dt_ioi[SubNr != 3], file = here("trimmed", "data_analysis_ioi.txt"), sep = ",")
fwrite(dt_kor[SubNr != 3], file = here("trimmed", "data_analysis_kor.txt"), sep = ",")
fwrite(dt_kot[SubNr != 3], file = here("trimmed", "data_analysis_kot.txt"), sep = ",")
fwrite(dt_vel[SubNr != 3], file = here("trimmed", "data_analysis_vel.txt"), sep = ",")
fwrite(dt_vel_diff[SubNr != 3], file = here("trimmed", "data_analysis_vel_diff.txt"), sep = ",")
```