### 2022-06-15 experiment-1/analysis/stats/variability.Rmd
start variabilityy analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Variability analysis"
output: html_notebook
---