### 2022-06-23 experiment-2/analysis/demographics/questionnaire.Rmd

---
title: 'Demographics: The Sound of Teaching Music 2'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of demographic information

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
```

# Questionnaire

```{r file, include = FALSE}
filename_q = "./questionnaire.csv"
```

```{r extract, echo = FALSE}
# read csv files
dt_2 <- fread(filename_q, header = T, sep = ",", dec = ".", na.string = "NA")
dt_2 <- dt_2[!21]

# change some colnames
colnames(dt_2)[2] <- "SubNr"
colnames(dt_2)[3] <- "Age"
colnames(dt_2)[4] <- "Gender"
colnames(dt_2)[7] <- "Handedness"
colnames(dt_2)[18] <- "PianoTotalPractice"
colnames(dt_2)[28] <- "TeachingMusic"

# extract the total number of years of training in piano (input manually based on the original csv file)
dt_2$PianoTotalPractice <- c(13, 15, 15, 16, 9, 10, 24, 8, 19, 6, 12, 10, 21, 23, 15, 22, 22, 24, 17, 15, 12)
dt_2$TeachingPianoYears <- c(NA, 4, NA, 2, NA, NA, NA, NA, 5, NA, NA, NA, 0.5, 9, 11, 2, NA, 8, 2, 1, 1)
dt_2$TeachingPiano <- "No"
dt_2[!is.na(TeachingPianoYears)]$TeachingPiano <- "Yes"

# change some characteristics
dt_2$Age <- as.numeric(dt_2$Age)
dt_2$PianoTotalPractice <- as.numeric(dt_2$PianoTotalPractice)

# exclude participants
dt_2_included <- dt_2[SubNr != 3]

print(dt_2_included)
```

## Descriptive stats for included participants
Note: A different questionnaire was used for only SubNr 21 and adjusted for necessary columns (but not others; see original csv for the precise information)

- Number of included participants: `r nrow(dt_2_included)` (mean age: `r round(mean(dt_2_included$Age), 2)`, sd: `r round(sd(dt_2_included$Age), 2)`)
- Male: `r nrow(dt_2_included[Gender == "Male"])` (mean age: `r round(mean(dt_2_included[Gender == "Male"]$Age), 2)`, sd: `r round(sd(dt_2_included[Gender == "Male"]$Age), 2)`)
- Female: `r nrow(dt_2_included[Gender == "Female"])` (mean age: `r round(mean(dt_2_included[Gender == "Female"]$Age), 2)`, sd: `r round(sd(dt_2_included[Gender == "Female"]$Age), 2)`)
- Average training years: `r round(mean(dt_2_included$PianoTotalPractice), 2)` years
- Handedness: Right (`r nrow(dt_2_included[Handedness == "Right"])`), Left (`r nrow(dt_2_included[Handedness == "Left"])`), Both (`r nrow(dt_2_included[Handedness == "Both"])`)
- Teaching experience (any instrument): Yes (`r nrow(dt_2_included[TeachingMusic == "Yes"])`), No (`r nrow(dt_2_included[TeachingMusic == "No"])`)
- Teaching experience (piano): Yes (`r nrow(dt_2_included[TeachingPiano == "Yes"])`, mean years of teaching: `r round( mean(dt_2_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`, sd: `r round(sd(dt_2_included[TeachingPiano == "Yes"]$TeachingPianoYears), 2)`), No (`r nrow(dt_2_included[TeachingPiano == "No"])`)

## Some descriptive questions
1. Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?

```{r 1, echo = FALSE}
data.table("Answer" = dt_2_included$`Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?`)
```

2. Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.

```{r 2, echo = FALSE}
data.table("Answer" = dt_2_included$`Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.`)
```

```{r export, include = FALSE}
knitr::purl("questionnaire.Rmd")
```