### 2022-06-27 experiment-1/analysis/stats/dynamics.Rmd
update additional analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Dynamics Analysis: The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of dynamics analysis.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Experimental Design{#design}
- Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: forte vs. piano / legato vs. staccato)
- Dependent variables: Key Velocity (KV), KV difference

1. Main predictions
- Participants will exaggerate forte and piano more while teaching.

2. Excluded participants

- SubNr 1: A deviated tempo
- SubNr 3: Experimental error
- SubNr 8: Experimental error
- SubNr 14: Experimental error
- SubNr 16: A deviated tempo

3. Stats packages
- [afex](https://github.com/singmann/afex) for ANOVAs
- [emmeans](https://github.com/rvlenth/emmeans) for posthoc comparisons

# Outline - Dynamics (Key Velocity)

## [Dynamics](#vel-dynamics)

### KV
- Individual results
  + [All](#vel_dyn_ind_all)
  + [Trial by trial](#vel_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#vel_dyn_group_all)
  + [Trial by trial](#vel_dyn_group_trial)

### KV-Diff
- Individual results
  + [All](#vel_diff_dyn_ind_all)
  + [Trial by trial](#vel_diff_dyn_ind_trial)

- Group results and stats (if available)
  + [All](#vel_diff_dyn_group_all)
  + [Trial by trial](#vel_diff_dyn_group_trial)

## [Articulation](#vel_articulation)

### KV
- Individual results
  + [All](#vel_art_ind_all)
  + [Trial by trial](#vel_art_ind_trial)

- Group results and stats (if available)
  + [All](#vel_art_group_all)
  + [Trial by trial](#vel_art_group_trial)

### KV-Diff
- Individual results
  + [All](#vel_diff_art_ind_all)
  + [Trial by trial](#vel_diff_art_ind_trial)

- Group results and stats (if available)
  + [All](#vel_diff_art_group_all)
  + [Trial by trial](#vel_diff_art_group_trial)

### [Sequence plot](#sequence)

- [Additional analysis based on reviewers' comments](#additional)

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
# statistics
if (!require("afex")) {install.packages("afex"); require("afex")}
if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}
```

```{r file, include = FALSE}
filename_vel = "../preprocessor/trimmed/data_analysis_vel.txt"
filename_vel_diff = "../preprocessor/trimmed/data_analysis_vel_diff.txt"
```

```{r extract, include = FALSE}
dt_vel <- fread(filename_vel, header = T, sep = ",", dec = ".") # read a trimmed txt
dt_vel_diff <- fread(filename_vel_diff, header = T, sep = ",", dec = ".") # read a trimmed txt

# SubNr as a factor
dt_vel$SubNr <- as.factor(dt_vel$SubNr)
dt_vel_diff$SubNr <- as.factor(dt_vel_diff$SubNr)

# Include only articulation
dt_vel_art <- dt_vel[Skill == "articulation"]
dt_vel_diff_art <- dt_vel_diff[Skill == "articulation"]

# Include only dynamics
dt_vel_dyn <- dt_vel[Skill == "dynamics"]
dt_vel_diff_dyn <- dt_vel_diff[Skill == "dynamics"]
```

# Key Velocity (KV) - Dynamics
## 1. All - individual{#vel_dyn_ind_all}
- Average KV for each participant.

```{r vel-dyn, echo = FALSE}
# For each individual
vel_dyn_trial <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]

vel_dyn <- vel_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_dyn
```

### Plot
#### Box plot
```{r vel-dyn-box,  echo = FALSE}
ggboxplot(vel_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Dynamics") +
  facet_grid(Subcomponent ~ .)
```

## 2. All - group{#vel_dyn_group_all}
- Average KV for each condition.
- Hypothesis: Participants will exaggerate forte and piano when teaching.

```{r vel-dyn-all, echo = FALSE}
# Group mean
vel_dyn_all <- vel_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_dyn_all
```

### Plot
#### Box plot
```{r vel-dyn-all-box, echo = FALSE}
vel_dyn_plot <- data.table(subject = vel_dyn[Condition == "teaching"]$SubNr, teaching = vel_dyn[Condition == "teaching"]$Mean, performing = vel_dyn[Condition == "performing"]$Mean, subcomponent = vel_dyn[Condition == "teaching"]$Subcomponent)

ggpaired(vel_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Dynamics")
```

### Statistics
#### Two-way ANOVA (aov_car)
```{r anova-dyn-vel-1, echo = FALSE}
vel_dyn_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_dyn)
summary(vel_dyn_aov)
```

#### Two-way ANOVA (aov)
```{r anova-dyn-vel-2, echo = FALSE}
vel_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_dyn)
summary(vel_dyn_aov_2)
```

#### Post-hoc (emmeans)
```{r post-hoc-1, echo = FALSE}
vel_dyn_posthoc <- test(pairs(emmeans(vel_dyn_aov,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
vel_dyn_posthoc
```

## 3. Trial by trial - individual{#vel_dyn_ind_trial}
- Average KV for each trial for each participant.

```{r vel-dyn-trial, echo = FALSE}
# For each individual
vel_dyn_trial
```

### Plot
#### Line graph
```{r vel-dyn-trial-line-forte, echo = FALSE, fig.height = 4}
ggline(dt_vel_dyn[Subcomponent == "Forte", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-dyn-trial-line-piano, echo = FALSE, fig.height = 4}
ggline(dt_vel_dyn[Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
  scale_x_continuous(breaks = seq(1,8,1))
```

## 4. Trial by trial - group{#vel_dyn_group_trial}
- Average KV for each trial for each condition.

```{r vel-dyn-trial-all, echo = FALSE}
# Group mean
vel_dyn_trial_all <- vel_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean.

```{r vel-dyn-trial-all-line-forte, echo = FALSE}
ggline(vel_dyn_trial[Subcomponent == "Forte"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Dynamics/Forte") +
    scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-dyn-trial-all-line-piano, echo = FALSE}
ggline(vel_dyn_trial[Subcomponent == "Piano"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Dynamics/Piano") +
    scale_x_continuous(breaks = seq(1,8,1))
```

# Key Velocity Difference (Difference) - Dynamics
## 1. All - individual{#vel_diff_dyn_ind_all}
- Average KV-Diff for each participant.

```{r vel-diff-dyn, echo = FALSE}
# For each individual
vel_diff_dyn_trial <- dt_vel_diff_dyn[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]

vel_diff_dyn <- vel_diff_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_diff_dyn
```

### Plot
#### Box plot
```{r vel-diff-dyn-box,  echo = FALSE}
ggboxplot(vel_diff_dyn_trial, "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Dynamics") +
  facet_grid(Subcomponent ~ .)
```

## 2. All - group{#vel_diff_dyn_group_all}
- Average KV-Diff for each condition.

```{r vel-diff-dyn-all, echo = FALSE}
# Group mean
vel_diff_dyn_all <- vel_diff_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_diff_dyn_all
```

### Plot
#### Box plot
```{r vel-diff-dyn-all-box, echo = FALSE}
vel_diff_dyn_plot <- data.table(subject = vel_diff_dyn[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn[Condition == "teaching"]$Mean, performing = vel_diff_dyn[Condition == "performing"]$Mean, subcomponent = vel_diff_dyn[Condition == "teaching"]$Subcomponent)

ggpaired(vel_diff_dyn_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Difference", title = "KV-Diff: Dynamics")
```

### Statistics
#### Two-way ANOVA (aov_car)
```{r anova-dyn-vel-diff-1, echo = FALSE}
vel_diff_dyn_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn)
summary(vel_diff_dyn_aov)
```

#### Two-way ANOVA (aov)
```{r anova-dyn-vel-diff-2, echo = FALSE}
vel_diff_dyn_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn)
summary(vel_diff_dyn_aov_2)
```

#### Post-hoc (emmeans)
```{r post-hoc-2, echo = FALSE}
vel_diff_dyn_posthoc <- test(pairs(emmeans(vel_diff_dyn_aov,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
vel_diff_dyn_posthoc
```

## 3. Trial by trial - individual{#vel_diff_dyn_ind_trial}
- Average KV-Diff for each trial for each participant.

```{r vel-diff-dyn-trial, echo = FALSE}
# For each individual
vel_diff_dyn_trial
```

### Plot
#### Line graph
```{r vel-diff-dyn-trial-line-ftop, echo = FALSE, fig.height = 4}
ggline(dt_vel_diff_dyn[Subcomponent == "FtoP", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-diff-dyn-trial-line-ptof, echo = FALSE, fig.height = 4}
ggline(dt_vel_diff_dyn[Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
  scale_x_continuous(breaks = seq(1,8,1))
```

## 4. Trial by trial - group{#vel_diff_dyn_group_trial}
- Average KV-Diff for each trial for each condition.

```{r vel-diff-dyn-trial-all, echo = FALSE}
# Group mean
vel_diff_dyn_trial_all <- vel_diff_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_diff_dyn_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean.

```{r vel-diff-dyn-trial-all-line-ftop, echo = FALSE}
ggline(vel_diff_dyn_trial[Subcomponent == "FtoP"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Dynamics/Forte to Piano") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-diff-dyn-trial-all-line-ptof, echo = FALSE}
ggline(vel_diff_dyn_trial[Subcomponent == "PtoF"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Dynamics/Piano to Forte") +
  scale_x_continuous(breaks = seq(1,8,1))
```

# Key Velocity (KV) - Articulation
## 1. All - individual{#vel_art_ind_all}
- Average KV for each participant.

```{r vel-art, echo = FALSE}
# For each individual
vel_art <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_art
```

### Plot
#### Box plot
```{r vel-art-box,  echo = FALSE}
ggboxplot(dt_vel_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "KV (0-127)", title = "KV: Articulation") +
  facet_grid(Subcomponent ~ .)
```

## 2. All - group{#vel_art_group_all}
- Average KV for each condition.

```{r vel-art-all, echo = FALSE}
# Group mean
vel_art_all <- vel_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_art_all
```

### Plot
#### Box plot
```{r vel-art-all-box, echo = FALSE}
vel_art_plot <- data.table(subject = vel_art[Condition == "teaching"]$SubNr, teaching = vel_art[Condition == "teaching"]$Mean, performing = vel_art[Condition == "performing"]$Mean, subcomponent = vel_art[Condition == "teaching"]$Subcomponent)

ggpaired(vel_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV: Articulation")
```

### Statistics
#### Two-way ANOVA (aov_car)
```{r anova-art-vel-1, echo = FALSE}
vel_art_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_art)
summary(vel_art_aov)
```

#### Two-way ANOVA (aov)
```{r anova-art-vel-2, echo = FALSE}
vel_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), data = vel_art)
summary(vel_art_aov_2)
```

#### Post-hoc (emmeans)
```{r post-hoc-3, echo = FALSE}
vel_art_posthoc <- test(pairs(emmeans(vel_art_aov,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
vel_art_posthoc
```

## 3. Trial by trial - individual{#vel_art_ind_trial}
- Average KV for each trial for each participant.

```{r vel-art-trial, echo = FALSE}
# For each individual
vel_art_trial <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
vel_art_trial
```

### Plot
#### Line graph
```{r vel-art-trial-line-legato, echo = FALSE, fig.height = 4}
ggline(dt_vel_art[Subcomponent == "Legato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Legato") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-art-trial-line-staccato, echo = FALSE, fig.height = 4}
ggline(dt_vel_art[Subcomponent == "Staccato", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") +
  scale_x_continuous(breaks = seq(1,8,1))
```

## 4. Trial by trial - group{#vel_art_group_trial}
- Average KV for each trial for each condition.

```{r vel-art-trial-all, echo = FALSE}
# Group mean
vel_art_trial_all <- vel_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean.

```{r vel-art-trial-all-line-legato, echo = FALSE}
ggline(vel_art_trial[Subcomponent == "Legato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Articulation/Legato") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-art-trial-all-line-staccato, echo = FALSE}
ggline(vel_art_trial[Subcomponent == "Staccato"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "KV (0-127)", title = "KV: Articulation/Staccato") +
  scale_x_continuous(breaks = seq(1,8,1))
```

# Key Velocity Difference (Difference) - Articulation{#vel_articulation}
## 1. All - individual{#vel_diff_art_ind_all}
- Average KV-Diff for each participant.

```{r vel-diff-art, echo = FALSE}
# For each individual
vel_diff_art <- dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent)]
vel_diff_art
```

### Plot
#### Box plot
```{r vel-diff-art-box,  echo = FALSE}
ggboxplot(dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent, BlockNr, TrialNr)], "SubNr", "Mean", color = "Condition", xlab = "SubNr", ylab = "Difference", title = "KV-Diff: Articulation") +
  facet_grid(Subcomponent ~ .)
```

## 2. All - group{#vel_diff_art_group_all}
- Average KV-Diff for each condition.

```{r vel-diff-art-all, echo = FALSE}
# Group mean
vel_diff_art_all <- vel_diff_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_diff_art_all
```

### Plot
#### Box plot
```{r vel-diff-art-all-box, echo = FALSE}
vel_diff_art_plot <- data.table(subject = vel_diff_art[Condition == "teaching"]$SubNr, teaching = vel_diff_art[Condition == "teaching"]$Mean, performing = vel_diff_art[Condition == "performing"]$Mean, subcomponent = vel_diff_art[Condition == "teaching"]$Subcomponent)

ggpaired(vel_diff_art_plot, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KV (0-127)", title = "KV-Diff: Articulation")
```

### Statistics
#### Two-way ANOVA (aov_car)
```{r anova-art-vel-diff-1, echo = FALSE}
vel_diff_art_aov <- aov_car(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_art)
summary(vel_diff_art_aov)
```

#### Two-way ANOVA (aov)
```{r anova-art-vel-diff-2, echo = FALSE}
vel_diff_art_aov_2 <- aov(Mean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_art)
summary(vel_diff_art_aov_2)
```

#### Post-hoc (emmeans)
```{r post-hoc-4, echo = FALSE}
vel_diff_art_posthoc <- test(pairs(emmeans(vel_diff_art_aov,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
vel_diff_art_posthoc
```

## 3. Trial by trial - individual{#vel_diff_art_ind_trial}
- Average KV-Diff for each trial for each participant.

```{r vel-diff-art-trial, echo = FALSE}
# For each individual
vel_diff_art_trial <- dt_vel_diff_art[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Subcomponent, TrialNr)]
vel_diff_art_trial
```

### Plot
#### Line graph
```{r vel-diff-artt-rial-line-ltos, echo = FALSE, fig.height = 4}
ggline(dt_vel_diff_art[Subcomponent == "LtoS", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") +
  scale_x_continuous(breaks = seq(1,8,1))
```

```{r vel-diff-art-trial-line-stol, echo = FALSE, fig.height = 4}
ggline(dt_vel_diff_art[Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)], x = "TrialNr", y = "Mean", facet.by = "SubNr", shape = "Condition", color = "Condition", xlab = "TrialNr", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") +
  scale_x_continuous(breaks = seq(1,8,1))
```

## 4. Trial by trial - group{#vel_diff_art_group_trial}
- Average KV-Diff for each trial for each condition.

```{r vel-diff-art-trial-all, echo = FALSE}
# Group mean
vel_diff_art_trial_all <- vel_diff_art_trial[Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, TrialNr)]
vel_diff_art_trial_all
```

### Plot
#### Line graph
- Error bars show SEM of the mean.

```{r vel-diff-art-trial-all-line-ltos, echo = FALSE}
ggline(vel_diff_art_trial[Subcomponent == "LtoS"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Articulation/Legato to Staccato") +
  scale_x_continuous(breaks=seq(1,8,1))
```

```{r vel-diff-art-trial-all-line-stol, echo = FALSE}
ggline(vel_diff_art_trial[Subcomponent == "StoL"], x = "TrialNr", y = "Mean", add = "mean_se", position = position_dodge(.2), shape = "Condition", color = "Condition", xlab = "Trial", ylab = "Difference", title = "KV-Diff: Articulation/Staccato to Legato") +
  scale_x_continuous(breaks=seq(1,8,1))
```

# Sequence plot{#sequence}
## Dynamics{#sequence_dyn}
```{r seq-dyn, fig.width = 9, fig.height = 3, echo = FALSE}
# For each individual
vel_dyn_seq <- dt_vel_dyn[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, RowNr)]

ggline(vel_dyn_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Note", ylab = "", title = "KV: Dynamics") +
  scale_x_continuous(breaks = seq(1,67,1))

vel_diff_dyn_seq <- dt_vel_diff_dyn[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Interval)]

ggline(vel_diff_dyn_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "", title = "KV-Diff: Dynamics") +
  scale_x_continuous(breaks = seq(1,66,1))
```

## Articulation{#sequence_art}
```{r seq-art, fig.width = 9, fig.height = 3, echo = FALSE}
# For each individual
vel_art_seq <- dt_vel_art[, .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(SubNr, Condition, Skill, RowNr)]

ggline(vel_art_seq, x = "RowNr", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Note", ylab = "", title = "KV: Articulation") +
  scale_x_continuous(breaks = seq(1,67,1))

vel_diff_art_seq <- dt_vel_diff_art[, .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(SubNr, Condition, Skill, Interval)]

ggline(vel_diff_art_seq, x = "Interval", y = "Mean", add = "mean_se", position = position_dodge(.2), linetype = "Condition", shape = "Condition", color = "Condition", xlab = "Interval", ylab = "", title = "KV-Diff: Articulation") +
  scale_x_continuous(breaks = seq(1,66,1))
```

# Additional analysis{#additional}
- Normalise KV difference by the range of KV value between forte and piano for each participant.

```{r, echo = FALSE}
# normalise diff by KV range
vel_diff_dyn_trial$NormalisedDiff <- 0
for (i in unique(vel_diff_dyn_trial$SubNr)){
  for (m in unique(vel_diff_dyn_trial$TrialNr)){

    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "teaching"]$Mean))

    vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$NormalisedDiff <- vel_diff_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean/(max(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean)-min(vel_dyn_trial[SubNr == i & TrialNr == m & Condition == "performing"]$Mean))

  }
}

# each participant
vel_diff_dyn_normalised <- vel_diff_dyn_trial[, .(N = .N, NormalisedMean = mean(NormalisedDiff), SD = sd(NormalisedDiff), SEM = sd(NormalisedDiff)/sqrt(.N)), by = .(SubNr, Condition, Skill, Subcomponent)]

# group
vel_diff_dyn_normalised_all <- vel_diff_dyn_normalised[, .(N = .N, Mean = mean(NormalisedMean), SD = sd(NormalisedMean), SEM = sd(NormalisedMean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent)]
vel_diff_dyn_normalised_all
```

## Plot
```{r, echo = FALSE}
vel_diff_dyn_plot_2 <- data.table(subject = vel_diff_dyn_normalised[Condition == "teaching"]$SubNr, teaching = vel_diff_dyn_normalised[Condition == "teaching"]$NormalisedMean, performing = vel_diff_dyn_normalised[Condition == "performing"]$NormalisedMean, subcomponent = vel_diff_dyn_normalised[Condition == "teaching"]$Subcomponent)

ggpaired(vel_diff_dyn_plot_2, cond = "performing", cond2 = "teaching", color = "condition", facet.by = "subcomponent", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "Normalised KV (0-127)", title = "KV-Diff: Dynamics")
```

## Stats
```{r, echo = FALSE}
# stats
vel_diff_dyn_aov_2 <- aov_car(NormalisedMean ~ Condition*Subcomponent + Error(SubNr/(Condition*Subcomponent)), vel_diff_dyn_normalised)
summary(vel_diff_dyn_aov_2)
```

```{r, echo = FALSE}
vel_diff_dyn_posthoc_2 <- test(pairs(emmeans(vel_diff_dyn_aov_2,~Condition|Subcomponent), adjust = "tukey"), joint = TRUE)
vel_diff_dyn_posthoc_2
```

```{r export, include = FALSE}
knitr::purl("dynamics.Rmd")
```