### 2022-08-30 experiment-1/analysis/teaching/teaching_art.Rmd
upload analysis between those who had teaching experience vs those who didn't
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: 'Articulation Analysis (Teaching vs. Non-Teaching): The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of articulation analysis  when comparing those who had teaching experiences and those who don't.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Experimental Design
- Independent variables: Condition (2 levels: teaching vs. performing) x Subcomponent (2 levels: legato vs. staccato / forte vs. piano) x Teaching (2 levels: Yes vs. No)
- Dependent variables: Key-Overlap Time (KOT), Key-Overlap Ratio (KOR)

1. Main predictions
- Participants will exaggerate legato and staccato more while teaching.

2. Excluded participants

- SubNr 1: A deviated tempo
- SubNr 3: Experimental error
- SubNr 8: Experimental error
- SubNr 14: Experimental error
- SubNr 16: A deviated tempo

<!-- 3. Stats packages -->
<!-- - [afex](https://github.com/singmann/afex) for ANOVAs -->
<!-- - [emmeans](https://github.com/rvlenth/emmeans) for posthoc comparisons -->

```{r setup, include = FALSE}
# packages
if (!require("here")) {install.packages("here"): require("here")}
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
# # statistics
# if (!require("afex")) {install.packages("afex"); require("afex")}
# if (!require("emmeans")) {install.packages("emmeans"); require("emmeans")}

# read info about teaching experience
source(here("experiment-1/analysis/demographics", "questionnaire.R"), chdir = TRUE)
```

```{r file, include = FALSE}
filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
filename_kot = "../preprocessor/trimmed/data_analysis_kot.txt"
```

```{r extract, include = FALSE}
dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed txt
dt_kot <- fread(filename_kot, header = T, sep = ",", dec = ".") # read a trimmed txt

# SubNr as a factor
dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)
dt_kot$SubNr <- as.factor(dt_kot$SubNr)

# For each individual, each block, each trial
ioi <- dt_ioi[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, BlockNr, TrialNr, Condition, Skill)]

# Calculate KOR for each interval (outliers were already determined based on KOT)
dt_kor <- data.table()
for (subnr in unique(dt_kot$SubNr)){
  for (block in unique(dt_kot[SubNr == subnr]$BlockNr)){
    for (trial in unique(dt_kot[SubNr == subnr & BlockNr == block]$TrialNr)){
      dt_current <- dt_kot[SubNr == subnr & BlockNr == block & TrialNr == trial]
      dt_current$KOR <- dt_current$KOT/ioi[SubNr == subnr & ioi$BlockNr == block & TrialNr == trial]$Mean
      dt_kor <- rbind(dt_kor, dt_current)
    }
  }
}

# Include only articulation
dt_kot_art <- dt_kot[Skill == "articulation"]
dt_kor_art <- dt_kor[Skill == "articulation"]

# Include only dynamics
dt_kot_dyn <- dt_kot[Skill == "dynamics"]
dt_kor_dyn <- dt_kor[Skill == "dynamics"]
```

# Key-Overlap Time (KOT) - Articulation
## 1. All - individual
- Average KOT for each participant.

```{r kot-art, echo = FALSE}
# For each individual
kot_art_trial <- dt_kot_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
setorder(kot_art_trial, "SubNr", "Condition", "Skill", "Subcomponent")

kot_art <- kot_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]

# add teaching info
kot_art$Teaching <- rep(dt_included$TeachingPiano, each = 4)
kot_art$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
kot_art
```

## 2. All - group
- Average KOT for each condition.

```{r kot-art-all, echo = FALSE}
# Group mean
kot_art_all <- kot_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, Teaching)]
kot_art_all
```

### Plot
#### Box plot
```{r kot-art-all-box, echo = FALSE}
kot_art_plot <- data.table(subject = kot_art[Condition == "teaching"]$SubNr, teaching = kot_art[Condition == "teaching"]$Mean, performing = kot_art[Condition == "performing"]$Mean, subcomponent = kot_art[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))

ggpaired(kot_art_plot[subcomponent == "Legato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation-Legato")

ggpaired(kot_art_plot[subcomponent == "Staccato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Articulation-Staccato")
```

# Key-Overlap Ratio (KOR) - Articulation
## 1. All - individual
- Average KOR for each participant.

```{r kor-art, echo = FALSE}
# For each individual
kor_art_trial <- dt_kor_art[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
setorder(kor_art_trial, "SubNr", "Condition", "Skill", "Subcomponent")

kor_art <- kor_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]

# add teaching info
kor_art$Teaching <- rep(dt_included$TeachingPiano, each = 4)
kor_art$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
kor_art
```

## 2. All - group
- Average KOR for each condition.

```{r kor-art-all, echo = FALSE}
# Group mean
kor_art_all <- kor_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, Teaching)]
kor_art_all
```

### Plot
#### Box plot
```{r kor-art-all-box, echo = FALSE}
kor_art_plot <- data.table(subject = kor_art[Condition == "teaching"]$SubNr, teaching = kor_art[Condition == "teaching"]$Mean, performing = kor_art[Condition == "performing"]$Mean, subcomponent = kor_art[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))

ggpaired(kor_art_plot[subcomponent == "Legato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation-Legato")

ggpaired(kor_art_plot[subcomponent == "Staccato"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOR", title = "KOR: Articulation-Staccato")
```

# Key-Overlap Time (KOT) - Dynamics
## 1. All - individual
- Average KOT for each participant.

```{r kot-dyn, echo = FALSE}
# For each individual
kot_dyn_trial <- dt_kot_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
setorder(kot_dyn_trial, "SubNr", "Condition", "Skill", "Subcomponent")

kot_dyn <- kot_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]

# add teaching info
kot_dyn$Teaching <- rep(dt_included$TeachingPiano, each = 4)
kot_dyn$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
kot_dyn
```

## 2. All - group
- Average KOT for each condition.

```{r kot-dyn-all, echo = FALSE}
# Group mean
kot_dyn_all <- kot_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, Teaching)]
kot_dyn_all
```

### Plot
#### Box plot
```{r kot-dyn-all-box, echo = FALSE}
kot_dyn_plot <- data.table(subject = kot_dyn[Condition == "teaching"]$SubNr, teaching = kot_dyn[Condition == "teaching"]$Mean, performing = kot_dyn[Condition == "performing"]$Mean, subcomponent = kot_dyn[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))

ggpaired(kot_dyn_plot[subcomponent == "Forte"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Dynamics-Forte")

ggpaired(kot_dyn_plot[subcomponent == "Piano"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOT (ms)", title = "KOT: Dynamics-Piano")
```

# Key-Overlap Ratio (KOR) - Dynamics
## 1. All - individual
- Average KOR for each participant.

```{r kor-dyn, echo = FALSE}
# For each individual
kor_dyn_trial <- dt_kor_dyn[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOR), SD = sd(KOR)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr, Subcomponent)]
setorder(kor_dyn_trial, "SubNr", "Condition", "Skill", "Subcomponent")

kor_dyn <- kor_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill, Subcomponent)]

# add teaching info
kor_dyn$Teaching <- rep(dt_included$TeachingPiano, each = 4)
kor_dyn$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 4)
kor_dyn
```

## 2. All - group
- Average KOR for each condition.

```{r kor-dyn-all, echo = FALSE}
# Group mean
kor_dyn_all <- kor_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N)), by = .(Condition, Skill, Subcomponent, Teaching)]
kor_dyn_all
```

### Plot
#### Box plot
```{r kor-dyn-all-box, echo = FALSE}
kor_dyn_plot <- data.table(subject = kor_dyn[Condition == "teaching"]$SubNr, teaching = kor_dyn[Condition == "teaching"]$Mean, performing = kor_dyn[Condition == "performing"]$Mean, subcomponent = kor_dyn[Condition == "teaching"]$Subcomponent, teachingexp = rep(dt_included$TeachingPiano, each = 2), teachingyears = rep(dt_included$TeachingPianoYears, each = 2))

ggpaired(kor_dyn_plot[subcomponent == "Forte"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics-Forte")

ggpaired(kor_dyn_plot[subcomponent == "Piano"], cond = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "KOR", title = "KOR: Dynamics-Piano")
```

```{r export, include = FALSE}
knitr::purl("teaching_art.Rmd")
```