### 2022-08-30 experiment-1/analysis/teaching/teaching_ioi.Rmd

---
title: 'Tempo Analysis (Teaching vs. Non-Teaching): The Sound of Teaching Music 1'
output:
  html_notebook: default
editor_options: 
  chunk_output_type: inline
---

Description: This is a summary of tempo analysis when comparing those who had teaching experiences and those who don't.

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

# Experimental Design
- Independent variables: Condition (2 levels: teaching vs. performing) x Technique (2 levels: articulation vs. dynamics) x Teaching (2 levels: Yes vs. No)
- Dependent variables: Interonset intervals (IOIs)

1. Main predictions
- Participants will play slower while teaching regardless of the techniques.

2. Excluded participants

- SubNr 1: A deviated tempo
- SubNr 3: Experimental error
- SubNr 8: Experimental error
- SubNr 14: Experimental error
- SubNr 16: A deviated tempo

<!-- 3. Stats packages -->
<!-- - [afex](https://github.com/singmann/afex) for ANOVAs -->
<!-- - [rstatix](https://github.com/kassambara/rstatix) for effect sizes -->

```{r setup, include = FALSE}
# packages
if (!require("here")) {install.packages("here"): require("here")}
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
# # statistics
# if (!require("afex")) {install.packages("afex"); require("afex")}
# if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}

# read info about teaching experience
source(here("experiment-1/analysis/demographics", "questionnaire.R"), chdir = TRUE)
```

```{r file, include = FALSE}
filename_ioi = "../preprocessor/trimmed/data_analysis_ioi.txt"
```

```{r extract, include = FALSE}
dt_ioi <- fread(filename_ioi, header = T, sep = ",", dec = ".") # read a trimmed csv

# SubNr as a factor
dt_ioi$SubNr <- as.factor(dt_ioi$SubNr)

# Include only articulation
dt_ioi_art <- dt_ioi[Skill == "articulation"]
# Include only dynamics
dt_ioi_dyn <- dt_ioi[Skill == "dynamics"]
```

# IOI-Articulation 
## 1. All - individual
- Average IOIs for each participant.

```{r ioi-art, echo = FALSE}
# For each individual
ioi_art_trial <- dt_ioi_art[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
setorder(ioi_art_trial, "SubNr", "Condition", "Skill")

ioi_art <- ioi_art_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]

# add teaching info
ioi_art$Teaching <- rep(dt_included$TeachingPiano, each = 2)
ioi_art$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 2)
ioi_art
```

## 2. All - group
- Average IOIs for each condition.

```{r ioi-art-all, echo = FALSE}
# Group mean
ioi_art_all <- ioi_art[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Condition, Skill, Teaching)]
ioi_art_all
```

### Plot
#### Box plot
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).

```{r ioi-art-all-box, echo = FALSE}
ioi_art_plot <- data.table(subject = ioi_art[Condition == "teaching"]$SubNr, teaching = ioi_art[Condition == "teaching"]$Mean, performing = ioi_art[Condition == "performing"]$Mean, teachingexp = dt_included$TeachingPiano, teachingyears = dt_included$TeachingPianoYears)

ggpaired(ioi_art_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Articulation") + geom_hline(yintercept = 188, linetype = "dashed")
```

# IOI-Dynamics
## 1. All - individual
- Average IOIs for each participant.

```{r ioi-dyn, echo = FALSE}
# For each individual
ioi_dyn_trial <- dt_ioi_dyn[, .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(SubNr, Condition, Skill, BlockNr, TrialNr)]
setorder(ioi_dyn_trial, "SubNr", "Condition", "Skill")

ioi_dyn <- ioi_dyn_trial[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(SubNr, Condition, Skill)]

# add teaching info
ioi_dyn$Teaching <- rep(dt_included$TeachingPiano, each = 2)
ioi_dyn$TeachingYears <- rep(dt_included$TeachingPianoYears, each = 2)
ioi_dyn
```

## 2. All - group
- Average IOIs for each condition.

```{r ioi-dyn-all, echo = FALSE}
# Group mean
ioi_dyn_all <- ioi_dyn[, .(N = .N, Mean = mean(Mean), SD = sd(Mean), SEM = sd(Mean)/sqrt(.N), Median = median(Mean), IQR = IQR(Mean)), by = .(Condition, Skill, Teaching)]
ioi_dyn_all
```

### Plot
#### Box plot
- A dashed line represents the tempo given by a metronome (80bpm / IOI-188ms).

```{r ioi-dyn-all-box, echo = FALSE}
ioi_dyn_plot <- data.table(subject = ioi_dyn[Condition == "teaching"]$SubNr, teaching = ioi_dyn[Condition == "teaching"]$Mean, performing = ioi_dyn[Condition == "performing"]$Mean, teachingexp = dt_included$TeachingPiano, teachingyears = dt_included$TeachingPianoYears)

ggpaired(ioi_dyn_plot, cond1 = "performing", cond2 = "teaching", color = "condition", facet.by = "teachingexp", line.size = 0.3, line.color = "gray", xlab = "Condition", ylab = "IOIs (ms)", title = "IOI: Dynamics") + geom_hline(yintercept = 188, linetype = "dashed")
```

```{r export, include = FALSE}
knitr::purl("teaching_ioi.Rmd")
```